### 2018-10-21 Compass_analysis.Rmd

---
title: "COMPASS_analysis"
author: "Taryn McLaughlin"
date: "9/23/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 3)
```

```{r p-value function, echo=TRUE}
library(dplyr)
pvals<-function(data, column){

    data<-filter(data, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))

    x<-split(data,data$TB) 
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))

    y<-split(data,data$SM) 
    B<-lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
    C<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB")))
    D<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB")))

    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
    )
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
             B$X$p.value,  C$X$p.value,   D$X$p.value)

    bonf_adj_pvals<-round(p.adjust(pvals, method="bonferroni"), 4)

    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)

    table<-(cbind(labels, round(pvals, 4), bonf_adj_pvals, fdr_adj_pvals))
    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
                       "Bonferonni Corrected p-values", "FDR Corrected p-vales")
    table
}
```

```{r COMPASS fits with Yerkes Donors, cache=TRUE, fig.height=10, fig.width=7}
library(dplyr)
library(COMPASS)
DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")
compass_data<-DF
compass_data<-rename(compass_data, 
    "IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count",
    "IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.1",
    "IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.2",
    "IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.3",
    "IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.4",
    "IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.5",
    "IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.6",
    "IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.7",
    "!IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.8",
    "!IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.9",
    "!IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.10",
    "!IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.11",
    "!IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.12",
    "!IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.13",
    "!IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.14",
    "!IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.15")

compass_data$TB<-factor(compass_data$TB, levels= c("N","HC","LTBI","TB"))
compass_data$SM<-factor(compass_data$SM, levels= c("N", "X", "SM"))
compass_data$Stim<-factor(compass_data$Stim, levels=c("UN", "PMA", "WCL", "PEP", "SEA", "SWAP"))
metadata<-dplyr::select(compass_data, c("ID","TB","SM","Stim"))
metadata$group<-paste(metadata$TB,metadata$SM)
counts<- as.matrix(select(compass_data, "IFNg&IL4&IL13&TNFa", 
         "IFNg&IL4&IL13&!TNFa", 
         "IFNg&IL4&!IL13&TNFa",
         "IFNg&IL4&!IL13&!TNFa",
         "IFNg&!IL4&IL13&TNFa",
         "IFNg&!IL4&IL13&!TNFa",
         "IFNg&!IL4&!IL13&TNFa",
         "IFNg&!IL4&!IL13&!TNFa",
         "!IFNg&IL4&IL13&TNFa",
         "!IFNg&IL4&IL13&!TNFa",
         "!IFNg&IL4&!IL13&TNFa",
         "!IFNg&IL4&!IL13&!TNFa",
         "!IFNg&!IL4&IL13&TNFa",
         "!IFNg&!IL4&IL13&!TNFa",
         "!IFNg&!IL4&!IL13&TNFa",
         "!IFNg&!IL4&!IL13&!TNFa"))

metadata$unique_id <- metadata$ID

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")

#WCL
n_wcl <- subset(counts, metadata$Stim =="WCL");
rownames(n_wcl) <- subset(metadata$unique_id,metadata$Stim=="WCL")
metadata_wcl<-subset(metadata,metadata$Stim=="WCL")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_wcl))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_wcl<-COMPASS::SimpleCOMPASS (n_s=n_wcl, n_u=n_u, meta=metadata_wcl, individual_id = "unique_id", iterations = 10000, replications = 8, verbose = TRUE)

#PEP
n_pep <- subset(counts, metadata$Stim =="PEP")
rownames(n_pep) <- subset(metadata$unique_id,metadata$Stim=="PEP")
metadata_pep<-subset(metadata,metadata$Stim=="PEP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pep))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pep<-COMPASS::SimpleCOMPASS (n_s=n_pep, n_u=n_u, meta=metadata_pep, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#PMA
n_pma <- subset(counts, metadata$Stim =="PMA")
rownames(n_pma) <- subset(metadata$unique_id,metadata$Stim=="PMA")
metadata_pma<-subset(metadata,metadata$Stim=="PMA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pma))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pma<-COMPASS::SimpleCOMPASS (n_s=n_pma, n_u=n_u, meta=metadata_pma, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SEA
n_sea <- subset(counts, metadata$Stim =="SEA")
rownames(n_sea) <- subset(metadata$unique_id,metadata$Stim=="SEA")
metadata_sea<-subset(metadata,metadata$Stim=="SEA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_sea))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_sea<-COMPASS::SimpleCOMPASS (n_s=n_sea, n_u=n_u, meta=metadata_sea, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SWAP
n_swap<- subset(counts, metadata$Stim =="SWAP")
rownames(n_swap) <- subset(metadata$unique_id,metadata$Stim=="SWAP")
metadata_swap<-subset(metadata,metadata$Stim=="SWAP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_swap))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_swap<-COMPASS::SimpleCOMPASS (n_s=n_swap, n_u=n_u, meta=metadata_swap, individual_id = "unique_id", iterations = 10000, 
                                  replications = 8, verbose = TRUE)
```

```{r COMPASS heatmaps yerkes, error =TRUE}
library(COMPASS)
plot(fit_pma, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"), 
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_wcl, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"), 
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_pep, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"), 
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
```

```{r COMPASS scores yerkes, error=TRUE}
library(COMPASS)
library(ggplot2)
library(knitr)
library(dplyr)

####### PMA
scores_pma<-COMPASS::scores(fit_pma)
g<-ggplot(scores_pma, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Functionality Score",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")

g<-ggplot(scores_pma, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_pma, "FS") , caption = "Non Specific CD4+ T cell Responses")
kable(pvals(scores_pma, "PFS"))

#### WHOLE CELL LYSATE
scores_wcl<-COMPASS::scores(fit_wcl)
g<-ggplot(scores_wcl, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
g<-ggplot(scores_wcl, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")
kable(pvals(scores_wcl, "FS"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(scores_wcl, "PFS"))

######## PEPTIDE
scores_pep<-COMPASS::scores(fit_pep)
g<-ggplot(scores_pep, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4 Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="Disease Status",
         y="Functionality Score")
g<-ggplot(scores_pep, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4 Responses", 
         subtitle="Polyfunctionality Score",
         caption="",
         x="Disease Status",
         y="Polyunctionality Score")

kable(pvals(scores_pep, "FS"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(scores_pep, "PFS"))
```

```{r COMPASS fits without Yerkes Donors, cache=TRUE, fig.height=10, fig.width=7}
library(dplyr)
library(COMPASS)
DF<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/CD4_count.csv")
compass_data<-filter(DF, TB!="N")
compass_data$SM<-factor(compass_data$SM, levels=c("X","SM"))
compass_data$TB<-factor(compass_data$TB, levels=c("HC","LTBI","TB"))
compass_data<-rename(compass_data, 
    "IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count",
    "IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.1",
    "IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.2",
    "IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.3",
    "IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.4",
    "IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.5",
    "IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.6",
    "IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.7",
    "!IFNg&IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.8",
    "!IFNg&IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.9",
    "!IFNg&IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.10",
    "!IFNg&IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.11",
    "!IFNg&!IL4&IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.12",
    "!IFNg&!IL4&IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.13",
    "!IFNg&!IL4&!IL13&TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.14",
    "!IFNg&!IL4&!IL13&!TNFa" = "Singlets.Lymphocytes.LD..CD3..CD4..G.4.13.T..Count.15")

compass_data$Stim<-factor(compass_data$Stim, levels=c("UN", "PMA", "WCL", "PEP", "SEA", "SWAP"))
metadata<-dplyr::select(compass_data, c("ID","TB","SM","Stim"))
metadata$group<-paste(metadata$TB,metadata$SM)
counts<- as.matrix(select(compass_data, "IFNg&IL4&IL13&TNFa", 
         "IFNg&IL4&IL13&!TNFa", 
         "IFNg&IL4&!IL13&TNFa",
         "IFNg&IL4&!IL13&!TNFa",
         "IFNg&!IL4&IL13&TNFa",
         "IFNg&!IL4&IL13&!TNFa",
         "IFNg&!IL4&!IL13&TNFa",
         "IFNg&!IL4&!IL13&!TNFa",
         "!IFNg&IL4&IL13&TNFa",
         "!IFNg&IL4&IL13&!TNFa",
         "!IFNg&IL4&!IL13&TNFa",
         "!IFNg&IL4&!IL13&!TNFa",
         "!IFNg&!IL4&IL13&TNFa",
         "!IFNg&!IL4&IL13&!TNFa",
         "!IFNg&!IL4&!IL13&TNFa",
         "!IFNg&!IL4&!IL13&!TNFa"))

metadata$unique_id <- metadata$ID

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")

#WCL
n_wcl <- subset(counts, metadata$Stim =="WCL");
rownames(n_wcl) <- subset(metadata$unique_id,metadata$Stim=="WCL")
metadata_wcl<-subset(metadata,metadata$Stim=="WCL")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_wcl))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_wcl_noN<-COMPASS::SimpleCOMPASS (n_s=n_wcl, n_u=n_u, meta=metadata_wcl, individual_id = "unique_id", iterations = 10000, replications = 8, verbose = TRUE)

#PEP
n_pep <- subset(counts, metadata$Stim =="PEP")
rownames(n_pep) <- subset(metadata$unique_id,metadata$Stim=="PEP")
metadata_pep<-subset(metadata,metadata$Stim=="PEP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pep))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pep_noN<-COMPASS::SimpleCOMPASS (n_s=n_pep, n_u=n_u, meta=metadata_pep, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#PMA
n_pma <- subset(counts, metadata$Stim =="PMA")
rownames(n_pma) <- subset(metadata$unique_id,metadata$Stim=="PMA")
metadata_pma<-subset(metadata,metadata$Stim=="PMA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_pma))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_pma_noN<-COMPASS::SimpleCOMPASS (n_s=n_pma, n_u=n_u, meta=metadata_pma, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SEA
n_sea <- subset(counts, metadata$Stim =="SEA")
rownames(n_sea) <- subset(metadata$unique_id,metadata$Stim=="SEA")
metadata_sea<-subset(metadata,metadata$Stim=="SEA")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_sea))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_sea_noN<-COMPASS::SimpleCOMPASS (n_s=n_sea, n_u=n_u, meta=metadata_sea, individual_id = "unique_id", iterations = 10000, 
                                 replications = 8, verbose = TRUE)

#SWAP
n_swap<- subset(counts, metadata$Stim =="SWAP")
rownames(n_swap) <- subset(metadata$unique_id,metadata$Stim=="SWAP")
metadata_swap<-subset(metadata,metadata$Stim=="SWAP")

n_u<-subset(counts,metadata$Stim =="UN")
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN")
inter<-intersect(rownames(n_u),rownames(n_swap))
n_u<-subset(counts,metadata$Stim =="UN" & metadata$ID %in% inter)
rownames(n_u) <- subset(metadata$unique_id,metadata$Stim=="UN" & metadata$ID %in% inter)

fit_swap_noN<-COMPASS::SimpleCOMPASS (n_s=n_swap, n_u=n_u, meta=metadata_swap, individual_id = "unique_id", iterations = 10000, 
                                  replications = 8, verbose = TRUE)
```

```{r COMPASS heatmaps, error =TRUE}
library(COMPASS)
plot(fit_pma_noN, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"),
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_wcl_noN, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"),
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
plot(fit_pep_noN, order_by_max_functionality = FALSE, markers=c("IL13","IL4","IFNg","TNFa"),
     row_annotation=c("TB","SM"), border_color=NA, fontsize=14, threshold=0)
```

```{r COMPASS scores, error =TRUE}
library(COMPASS)
library(ggplot2)
library(knitr)

####### PMA
scores_pma_noN<-COMPASS::scores(fit_pma_noN)
g<-ggplot(scores_pma_noN, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Functionality Score",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
g<-ggplot(scores_pma_noN, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses",
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_pma_noN, "FS"), caption = "Non-Specific CD4+ T cell Responses")
kable(pvals(scores_pma_noN, "PFS"))

#### WHOLE CELL LYSATE
scores_wcl_noN<-COMPASS::scores(fit_wcl_noN)
g<-ggplot(scores_wcl_noN, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="Disease Status",
         y="Functionality Score")
g<-ggplot(scores_wcl_noN, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="Polyfunctionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_wcl_noN, "FS"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(scores_wcl_noN, "PFS"))

######## PEPTIDE
scores_pep_noN<-COMPASS::scores(fit_pep_noN)
g<-ggplot(scores_pep_noN, aes(x=SM, y=FS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4+ T cell Responses", 
         subtitle="Functionality Scores",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
g<-ggplot(scores_pep_noN, aes(x=SM, y=PFS, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(~TB, scale="free")+
    labs(title="Peptide Specific CD4+ T cell Responses", 
         subtitle="Polyfunctionality Score",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyunctionality Score")

kable(pvals(scores_pep_noN, "FS"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
kable(pvals(scores_pep_noN, "PFS"))

```