### 2018-10-21 ICS_CD4.Rmd

---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
library(dplyr)
library(reshape2)
Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)

melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  "x_4_13_x","x_4_x_x", "x_x_13_x"))
melt$TB<-factor(melt$TB, levels=c("N","HC","LTBI","TB"))
melt$SM<-factor(melt$SM, levels=c("N","X","SM"))
melt<-filter(melt, !is.na(label))

data<-filter(melt, TB %in% c("HC", "N") & Stim=="PMA")

g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#67001f"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(Stim~TB, scale="free")

    scale_color_manual(labels = c("SM(-)", "SM(+)"), values = Palette) +
    scale_x_discrete(breaks=c("X","SM"),
                     labels=c("SM(-)", "SM(+)"))+
    theme_bw()+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_wrap(~TB,nrow = 1, scale="free_x")+
    labs(title="Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and Helminth Status",
         y="Frequency of Cytokine+ CD4+ T cells")

```{r}
#Antigen specific responses using total cytokine
Cyt<-read.csv("Cyt.csv")
TH1<-read.csv("TH1.csv")
TH2<-read.csv("TH2.csv")
TH1.2<-read.csv("TH1-2.csv")

Split<- strsplit(as.character(Cyt$Sample)," ")
Cyt$Sample = as.factor(sapply(Split,"[",2))
TH1$Sample = as.factor(sapply(Split,"[",2))
TH2$Sample = as.factor(sapply(Split,"[",2))
TH1.2$Sample = as.factor(sapply(Split,"[",2))

AG<-AG[!(AG$Sample=="NK2037_TB_HEL_T_PMA_020.fcs")]
AG<-AG[!(AG$Sample=="NK2126_HC_X_SEA_005.fcs")]
AG<-AG[!(AG$Sample=="NK2450_HC_SM_SEA_047.fcs")]
AG<-AG[!(AG$Sample=="NK2402_LTBI_SM_SWAP_024.fcs")]
AG<-AG[!(AG$Sample=="NK2342_TB_X_SWAP_006.fcs")]
AG<-AG[!(AG$Sample=="NK2136_TB_X_WCL_003.fcs")]
AG<-AG[!(AG$Sample=="NK2168_HC_SM_WCL_040.fcs")]
AG<-AG[!(AG$Sample=="NK2088_LTBI_HEL_H_PEP_010.fcs")]

AG<-as.data.table(AG)

TH1<-merge(AG,TH1,by="Sample")
TH2<-merge(AG,TH2,by="Sample")
TH1.2<-merge(AG,TH1.2,by="Sample")
Cyt<-merge(AG,Cyt,by="Sample")

Cyt<-Cyt[,c(1:5,21:25,29,30,35:51)]
TH1<-TH1[,c(1:5,21,24,26,31,35:51)]
TH2<-TH2[,c(1:5,22:23,28,32,35:51)]
TH1.2<-TH1.2[,c(1:5,21:24,27,33,35:51)]

```

```{r}

```

Prep for Stacked Bars PMA as example
```{r}
df<-WCL[,c(4,5,27:34)]
df$TH1.Portion<-(df$TH1.Count/df$Cyt.Count)
df$TH2.Portion<-(df$TH2.Count/df$Cyt.Count)
df$TH1.2.Portion<-(df$TH1.2.Count/df$Cyt.Count)
df$TH1.Port<-(df$TH1.Freq/df$Cyt.Freq)
df$TH2.Port<-(df$TH2.Freq/df$Cyt.Freq)
df$TH1.2.Port<-(df$TH1.2.Freq/df$Cyt.Freq)

#unscaled frequencies
df1<-df[,c(1,2,6,8,10)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
df3$variable<-factor(df3$variable, levels=c('TH2.Freq','TH1.2.Freq','TH1.Freq',ordered=TRUE))
df3$Label <- paste(df3$TB, df3$SM, sep='_')

#adjusting error HC_X
df3$upper[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]<-
  df3$median[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]+
  df3$median[df3$variable=="TH1.Freq" & df3$Label=="HC_X"]+
  df3$error[df3$variable=="TH1.2.Freq" & df3$Label=="HC_X"]

```

Boolean Cytokines with PMA as an example
```{r}
# to make the cytokine frequency graphs example with PMA
library(dplyr)
library(reshape2)
Boolean<-select(data, Donor, TB, SM, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T,G_4_x_x, G_x_13_x,x_4_x_T,x_x_13_T, G_x_x_T , G_x_x_x,x_x_x_T,  x_4_13_x,x_4_x_x, x_x_13_x)

melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  "x_4_13_x","x_4_x_x", "x_x_13_x"))

HC<-subset(melt, TB %in% "HC")
LTBI<-subset(melt, TB %in% "LTBI")
ATB<-subset(melt, TB %in% "TB")

```

Surface Markers on CD4
```{r}
df1<-CD4[,c(7:9,19,20)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
df3$Label <- paste(df3$TB, df3$SM, sep='_')

df3$variable<-factor(df3$variable, levels=c('CCR4.P_CXCR3.P','CCR4.N_CXCR3.P','CCR4.P_CXCR3.N', ordered=TRUE))
```

Transcription Factors on CD4
```{r}
df1<-data[,c(22:24,4,5)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
df3$Label <- paste(df3$TB, df3$SM, sep='_')
df3$variable<-factor(df3$variable, levels=c("TH1_GATA3.P_Tbet.P",'TH1_GATA3.P_Tbet.N','TH1_GATA3.N_Tbet.P', ordered=TRUE))
``` 
Surface Markers on AG specific
```{r}
df<-TH1[TH1$Stim=="WCL"]
df1<-df[,c(9:11,20,21)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
df3$Label <- paste(df3$TB, df3$SM, sep='_')
df3$variable<-factor(df3$variable, levels=c('CCR4.P_CXCR3.P','CCR4.N_CXCR3.P','CCR4.P_CXCR3.N', ordered=TRUE))
```

Transcription Factors on AG specific
```{r}
df<-TH1[TH1$Stim=="WCL"]
df1<-df[,c(15:17,20,21)]
df2<-melt(df1,id.vars=c("TB","SM"))
df3<-as.data.frame(
  ddply(df2, c("TB", "SM", "variable"), summarise,
        median = median(value), sd = sd(value),
        error = qnorm(0.975)*sd(value)/sqrt(length(value))))
df3$upper<-(df3$median+df3$error)
df3$lower<-(df3$median-df3$error)

#you have to do these to get the bars to be right
# (grey, neither, top), (cornflower, mix), (orchid, type 2), (navy, type 1, bottom)
df3$Label <- paste(df3$TB, df3$SM, sep='_')
df3$variable<-factor(df3$variable, levels=c("GATA3.P_Tbet.P",'GATA3.P_Tbet.N','GATA3.N_Tbet.P', ordered=TRUE))
```

Marker Co-expression
```{r}
CD4_melt<-melt(data,id.vars=c("TB","SM"))
CD4_melt$value<-as.numeric(CD4_melt$value)
CD4_melt<-CD4_melt[!(is.na(CD4_melt$value))]
CD4_stats<-as.data.frame(
  ddply(CD4_melt, c("TB", "SM", "variable"), summarise,
        mean = mean(value), sd = sd(value),
        sem = sd(value)/sqrt(length(value))))

TF<-subset(CD4_melt, variable %in% c("GATA3.N_Tbet.P", "GATA3.P_Tbet.N", "GATA3.P_Tbet.P"))
tf<-subset(CD4_stats, variable %in% c("GATA3.N_Tbet.P", "GATA3.P_Tbet.N", "GATA3.P_Tbet.P"))
CC<-subset(CD4_melt, variable %in% c("CCR4.N_CXCR3.P", "CCR4.P_CXCR3.N", "CCR4.P_CXCR3.P"))
cc<-subset(CD4_stats, variable %in% c("CCR4.N_CXCR3.P", "CCR4.P_CXCR3.N", "CCR4.P_CXCR3.P","CCR4.N_CXCR3.N"))
```

stats? sort of read this http://rcompanion.org/rcompanion/d_08.html
```{r}
data<-subset(CYT,CYT$Stim=="WCL")
data<-filter(TH1,(Stim=="WCL" & TH1.Sum>0))

x <- split(data, data$TB)
lapply(x, function(g) wilcox.test(g$Cyt.Freq~g$SM))
lapply(x, function(g) wilcox.test(g$TH1.Freq~g$SM))
lapply(x, function(g) wilcox.test(g$TH1.2.Freq~g$SM))
lapply(x, function(g) wilcox.test(g$TH2.Freq~g$SM))

y <- split(data, data$SM)
lapply(y, function(g) kruskal.test(g$Cyt.Freq~g$TB))
lapply(y, function(g) kruskal.test(g$TH1.Freq~g$TB))
lapply(y, function(g) kruskal.test(g$TH1.2.Freq~g$TB))
lapply(y, function(g) kruskal.test(g$TH2.Freq~g$TB))

data<-filter(TH1,(Stim=="WCL" & TH1.Sum>0))
x <- split(data, data$TB)
lapply(x, function(g) wilcox.test(g$TH1_Tbet~g$SM))
lapply(x, function(g) wilcox.test(g$TH1_GATA3~g$SM))

y <- split(data, data$SM)
lapply(y, function(g) kruskal.test(g$TH1_Tbet~g$TB))
lapply(y, function(g) kruskal.test(g$TH1_GATA3~g$TB))
```