### 2018-10-21 Stock Code.Rmd

---
title: "Stock Code"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#concatenating tables- "fill" includes columns that are different between the two
OG<-rbind.fill(OG.5,OG.8)

#renaming columns
setnames(OG, 
         old=c("Sample","Singlets.Lymphocytes.LD..CD3..Q1..CD4_...CD8..Count") ,
         new=c("StudyID","CD8.Pos"))

#splitting values of a variable by some parameter (only works for strings: factors & characters)
split = strsplit(as.character(OG$StudyID),":")

#create a new column with the a values from the split
OG$Donor = as.factor(sapply(splStudyID,"[",2))

# REORDER A FACTOR VARIABLE
OG$TEST = factor(OG$Antigen,levels=c("UN","PPD","WCL","PEP","SEB"))

#delete entire rows based on the value in a cell- must be a data.table
OG<- OG[!(Donor=="NA")]

#Melt your data table- id.vars are the variables which you want to "melt" by
#can melt a data.table or a data.frame 
melted<-melt(OG, id.vars=c("Antigen", "RE"))
#for some reason it stores all the values as character so coerce them to numeric to run stats
melted$value<-as.numeric(as.character(melted$value))

#summary statistics spit out as a data fram
stats<-as.data.frame(
  ddply(DATA, c("Antigen", "RE", "variable"), summarise,
        mean = mean(value), sd = sd(value),
        sem = sd(value)/sqrt(length(value))))

# trying to mass turn data to numeric from factor
jfhuigls <- data.frame(lapply(OG.RE, function(x) as.numeric(as.character(x))))
melted$value<-as.numeric(as.character(melted$value))

#do a T-test on two subsets of the data
t.test(subset(melted$value,melted$Antigen1 == "UN" & melted$variable=="IFN"),
       subset(melted$value,melted$Antigen1 == "SEB" & melted$variable=="IFN"))

#summary statistics for graphing purposes--> makes a function which can be applied to any data you want
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)

  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }

  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )

  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))

  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult

  return(datac)
}

# how to apply the summary statistic function
Dat<-summarySE(OG, measurevar="CD4_OG.L", groupvars=c("Antigen","Time"))

#and then graph it
pd <- position_dodge(0.1) # move error bars .05 to the left and right to avoid overlap
ggplot(Dat, aes(x=Antigen, y=CD4_OG.L, colour=Time)) + 
  geom_errorbar(aes(ymin=CD4_OG.L-se, ymax=CD4_OG.L+se), colour="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=3) +
  xlab("Stimulation") +
  ylab("Frequency of Oregon Green Low CD4+ T cells") +
  ggtitle("Proliferation Restimulation Assay") +
  theme_bw()

#simplify graph and use 95% confidence interval instead of SEM
ggplot(tgc, aes(x=dose, y=len, colour=supp)) + 
  geom_errorbar(aes(ymin=len-ci, ymax=len+ci), width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd)

#change individual cells based on their content
levels(TB$TBAntigen)[levels(TB$TBAntigen)==">10"] <- "10"
ELISA$Result[ELISA$Result>500]<-500
HIV$HIVResult[HIV$HIVResult==1]<-as.character("POS")

#remove columns
TB<-TB[,c(2,4,5,9,11:15):=NULL]

#subset data to make it easier to work with
NKS<-subset(Final,Final$Study=="NKS")
TBRU<-subset(Final,Final$Study=="TRU")

## create new column with NA as value
TBRU[,"TB.Status"]<-NA

## for each row in the data table
## update new column with value based on other column
for (i in 1:nrow(TBRU)){
  if(TBRU$QualResult[i]=="1") {TBRU$TB.Status[i]<-as.character("LTBI")}
  else {TBRU$TB.Status[i]<-as.character("Healthy Control")}
}

for (i in 1:length(Final$HIV.Status)){
  Final$HIV.Status[i] <- ifelse(is.na(Final$HIVResult[i]), as.character(Final$HIV.STATUS[i]),
                                as.character(Final$HIVResult[i]))
}

#Aggregate data from the same donor in a data set and perform some function on it
##this makes a new data frame with the StudyID and the aggregated data
A = aggregate(AscarisPositive ~ StudyID, data = HEL, function(x) {

  if (1 %in% x) {
    return(1);
  } else {
    return(0);
  }
});

#merge two data sets. match rows by study IDs 
##but keep all rows regardless of whether the StudyID is in both sets
I<-merge(Disease,TB,by="StudyID",all=TRUE)

#Check for duplicated IDs
dups<-Final$StudyID[duplicated(Final$StudyID)]
#NK2081 NK2194 NK2269 NK2464 NK2667 NK2712 80008  80019

#find common StudyIDs and remove the non common StudyIDs from a data set
Inter<- intersect(HEL$StudyID,TB$StudyID)
CleanTB<-TB[TB$StudyID %in% Inter,];

#no clue figure these out
s<-split(Final, Final$HelmithPositive)
i<-interaction(Final$HelmithPositive, Final$QualResult)
S<-str(split(Final,list(Final$HelmithPositive, Final$QualResult), drop=TRUE))

#graph
qplot(test$SchistosomaIntensity,test$QFT, col=test$SchistosomaPositive.y)

#correlation test
# the [] provides a restriction on the data used in the correlation
cor.test((test$SchistosomaIntensity)[test$SchistosomaPositive==1],test$QFT[test$SchistosomaPositive==1])

# linear model
lm((test$SchistosomaIntensity)[test$SchistosomaPositive==1]~test$QFT[test$SchistosomaPositive==1])

#ANOVA
aov((test$SchistosomaIntensity)[test$SchistosomaPositive==1]~test$QFT[test$SchistosomaPositive==1])

###power for proportions
trans.arcsine <- function(x){
  asin(sign(x) * sqrt(abs(x)))
}
h=2*trans.arcsine(b/d)-2*trans.arcsine(a/c)
pwr.2p.test(h=h,n=NULL,sig.level=0.05,power=.8)

for (i in 1:length(Cyt$IFNg)) {
  Cyt$IFNg[i]<-ifelse(Cyt$Stim[i]=="UN",Cyt$IFNg[i],(Cyt$IFNg[i]-Cyt$IFNg[Cyt$Stim=="UN"]))
}

#PCA Google this https://www.r-bloggers.com/computing-and-visualizing-pca-in-r/
#http://www.statmethods.net/stats/frequencies.html

#graphing data http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/#Helper functions
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.