### 2019-03-27 Cytokine_Analysis.Rmd

---
title: "Cytokine_Analysis"
author: "Taryn McLaughlin"
date: "9/21/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)

source("/Users/tarynam/Desktop/tryna function.R")
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
data<-read.csv("CD4_Cytokine_clean.csv")
    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
    data$SM<-factor(data$SM, levels=c("N","X","SM"))
CD4<-data

data<-read.csv("CD8_Cytokine_clean.csv")
    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
    data$SM<-factor(data$SM, levels=c("N","X","SM"))
CD8<-data

data<-read.csv("CD4_MFI_clean.csv")
    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
    data$SM<-factor(data$SM, levels=c("N","X","SM"))
MFI<-data

data<-read.csv("ICS_Compass_scores_complete.csv")
    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
    data$SM<-factor(data$SM, levels=c("N","X","SM"))
Compass<-data

data<-read.csv("GD_Cytokine_clean.csv")
    data$TB<-factor(data$TB, levels=c("N","HC","LTBI","TB"))
    data$SM<-factor(data$SM, levels=c("N","X","SM"))
GD<-data

data<-read.csv("ICS_Donor_Info_clean.csv")
data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
data$TB<-factor(data$TB, levels=c("Naïve","Healthy Control","LTBI","Active TB"))
data$SM<-gsub("US|Not US|Unknown", "Naïve", data$SM)
data$SM<-factor(data$SM, levels=c("Naïve","Negative","Schisto+"))
Donor<-data
```

```{r narrow down Yerkes Donors}
library(dplyr)
mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
HD<-filter(mimosa, grepl("HD", mimosa$Sample))%>%
    filter(Stim!="PMA")
HDscores<-aggregate(score~Donor, HD, FUN=sum)
HDkeep<-head(arrange(HDscores, score), 10)
#382 479 487 are BCG
HDremove<-HD$Donor[!(HD$Donor %in% HDkeep$Donor)]
CD4<-dplyr::filter(CD4, !(Donor %in% HDremove))
MFI<-dplyr::filter(MFI, !(Donor %in% HDremove))
Compass<-dplyr::filter(Compass, !(Donor %in% HDremove))
Donor<-dplyr::filter(Donor, !(Donor %in% HDremove))
```

```{r viability and cell counts}
Donor$TB[grepl("HD", Donor$StudyID)==TRUE]<-"Yerkes"
palette.four<-c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b")
ggplot(Donor, aes(x=SM, y=Viability, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="Viability", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Live Cells (%)")

ggplot(Donor, aes(x=SM, y=Cells.Vial, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="Cell Counts", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Cells/mL)")
```

#Statistical Analysis
I wrote a function called pvals_helminth which evaluates the relevant two sample two-tailed  p-values using the non-parametric Mann Whitney test (the command is wilcox.test in R). The function then corrects for multiple comparisons by controlling the False Discovery Rate, the expected proportion of false discoveries amongst the rejected hypotheses.

```{r p-value function, echo=TRUE}
# a function to evaluate differences in the expression of a marker (designated by the "column" argument) between disease groups
pvals_helminth<-function(data, column){

    x<-split(data,data$TB) #creates individual dataframes for each level of TB
    #does a wilcox test for the column of interest comparing SM- to SM+ within each level of TB
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"], exact=FALSE))

    y<-split(data,data$SM) #creates two dataframes for SM- and SM+
    #each lapply does a wilcox test on subsets these dataframes to compare 2 levels of TB
    #stores the output in a variable. So B will contain the comparison of HC to LTBI for both SM- and SM+
    B<-lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI"), exact=FALSE))
    C<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB"), exact=FALSE))
    D<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB"), exact=FALSE))

    #To keep track of all the p-values... these should match the order below
    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
    )

    #extract the p-values from all the stored wilcox tests
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
             B$X$p.value,  C$X$p.value,   D$X$p.value)

    #adjusts the p-values using the false discovery rate strategy
    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)

    #create a table with the output
    table<-(cbind(labels, round(pvals, 4), fdr_adj_pvals))
    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
                       "FDR Corrected")
    table #return the table
}
```

#Compass Graphs
```{r}
library(dplyr)
DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype =="CD4")
ggplot(DF, aes(x=SM, y=FS, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=4)+
    scale_alpha_manual(values=c(0.5,1))+
    scale_fill_manual(values = c("#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
      theme(plot.title = element_text(hjust = 0.5)) +
    scale_x_discrete(labels=c("SM-", "SM+"))+
    facet_grid(~TB, scales="free")+
    labs(title="WCL-Specific CD4+ T cell Response", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Functionality Score")
ggplot(DF, aes(x=SM, y=PFS, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="WCL-Specific CD4+ T cell Response", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Polyfunctionality Score")

DF2<-merge(DF, dplyr::select(Donor, Donor, QFT))
ggplot(DF2, aes(x=QFT, y=FS, col=SM))+geom_point()+facet_grid(~TB, scales = "free")+geom_smooth(method="lm")
```

#Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it incredibly cluttered looking, however I did allow outliers to display.
```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)

Boolean<-dplyr::select(CD4, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x, x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)%>%
    dplyr::filter(TB!="N")

melt<-melt(Boolean,id.vars=c("Donor","TB","SM", "Stim"))

#all the boolean plots
ggplot(melt, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(Stim~TB, scale="free")+
    labs(title="Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

#subset by stim
data<-filter(melt, Stim=="SWAP")
ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#009E73" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="fixed")+
    labs(title="SWAP-Specific Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

#3 separate booleans for TH1/2 TH1 and TH2
#TH1-2
data<-filter(melt, Stim=="SWAP")%>%
    filter(variable %in% c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T"))
ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1))+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    facet_grid(TB~., scale="fixed")+
    coord_cartesian(ylim = c(0,.005))

data<-filter(melt, Stim=="SWAP")%>%
    filter(variable %in% c("G_x_x_T" , "G_x_x_x","x_x_x_T"))
g<-ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    facet_grid(TB~., scale="fixed")+
    coord_cartesian(ylim = c(0,.2))

data<-filter(melt, Stim=="SWAP")%>%
    filter(variable %in% c("x_4_13_x","x_4_x_x", "x_x_13_x"))
g<-ggplot(data, aes(x=variable, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    facet_grid(TB~., scale="fixed")+
    coord_cartesian(ylim = c(0,.05))
```

#T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups.

```{r TH, echo=FALSE, fig.cap="Breakdown of Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH_Group_Explanation.png")
```

```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
TH<-dplyr::select(CD4, Donor, TB, SM, Stim, Cyt.Freq, Th1.Freq, TH1.2.Freq, TH2.Freq)

melt<-melt(TH,id.vars=c("TB","SM", "Stim"), measure.vars=c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq"))
melt$variable = factor(melt$variable,levels=c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq"))

##### PMA
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(variable~TB, scale="free")+
    labs(title="Non Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="PMA")
kable(pvals_helminth(data, "Th1.Freq"), caption = "Non Specific CD4+ T cell Responses")
kable(pvals_helminth(data, "TH2.Freq"))
kable(pvals_helminth(data, "TH1.2.Freq"))

##### WCL
data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(variable~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="WCL")
kable(pvals_helminth(data, "Th1.Freq"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
kable(pvals_helminth(data, "TH2.Freq"))
kable(pvals_helminth(data, "TH1.2.Freq"))

##### Peptide
data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(variable~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="PEP")
kable(pvals_helminth(data, "Th1.Freq"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals_helminth(data, "TH2.Freq"))
kable(pvals_helminth(data, "TH1.2.Freq"))

```
\newpage

#Individual Cytokines
These are the frequencies of cells producing individual cytokines without considering whether or not they are single positive, double positive etc. So any cell producing IL-4 is included in the IL-4 analysis.

```{r Individual Cytokines but not single producers}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
cyt<-select(Cytokine, Donor, TB, SM, Stim, IFNg, TNFa, IL4, IL13)

melt<-melt(cyt,id.vars=c("TB","SM", "Stim"), measure.vars=c("IFNg", "TNFa", "IL4", "IL13"))
melt$variable = factor(melt$variable,levels=c("IFNg", "TNFa", "IL4", "IL13"))

#### PMA
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="PMA")
kable(pvals_helminth(data, "IFNg"), caption = "Non Specific CD4 T cell Responses")
kable(pvals_helminth(data, "TNFa"))
kable(pvals_helminth(data, "IL4"))
kable(pvals_helminth(data, "IL13"))

#### WCL
data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="WCL")
kable(pvals_helminth(data, "IFNg"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
kable(pvals_helminth(data, "TNFa"))
kable(pvals_helminth(data, "IL4"))
kable(pvals_helminth(data, "IL13"))

#### Peptide
data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="PEP")
kable(pvals_helminth(data, "IFNg"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals_helminth(data, "TNFa"))
kable(pvals_helminth(data, "IL4"))
kable(pvals_helminth(data, "IL13"))

```

#True Single and Double Cytokine Positive CD4+ T Cells
These are the frequencies of cells truly ONLY producing:  
1. IFNg or TNFa or the combination of both AND NOT IL4 or IL13  
2. IL-4 or IL-13 or the combination of both AND NOT IFNg or TNFa  
These are the subgroups within the grouped TH1 vs TH2 lineage analysis. They are essentially the last 6 groups of the boolean data.

```{r Single and Double Producers Lin Spec graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
SP_lin<-rename(Cytokine, G_single =  G_x_x_x, T_single = x_x_x_T, IL4_single = x_4_x_x, IL13_single = x_x_13_x, G_T = G_x_x_T, IL4_IL13 = x_4_13_x) %>%
    select(Donor, TB, SM, Stim, G_single, T_single, IL4_single, IL13_single, G_T, IL4_IL13)

melt<-melt(SP_lin,id.vars=c("TB","SM", "Stim"), measure.vars=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))
melt$label = factor(melt$variable,levels=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))

#### PMA
# Th1 plots
data<-filter(melt, Stim=="PMA" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PMA" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="PMA")
kable(pvals_helminth(data, "G_single"), caption = "Non-Specific CD4+ T cell Responses")
kable(pvals_helminth(data, "T_single")) 
kable(pvals_helminth(data,"G_T"))
kable(pvals_helminth(data, "IL4_single"))
kable(pvals_helminth(data, "IL13_single")) 
kable(pvals_helminth(data, "IL4_IL13"))

#### WCL
# Th1 plots
data<-filter(melt, Stim=="WCL" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="WCL" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="WCL")
kable(pvals_helminth(data, "G_single"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals_helminth(data, "T_single")) 
kable(pvals_helminth(data,"G_T"))
kable(pvals_helminth(data, "IL4_single"))
kable(pvals_helminth(data, "IL13_single")) 
kable(pvals_helminth(data, "IL4_IL13"))

#### PEP
# Th1 plots
data<-filter(melt, Stim=="PEP" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PEP" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="PEP")
kable(pvals_helminth(data, "G_single"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals_helminth(data, "T_single")) 
kable(pvals_helminth(data,"G_T"))
kable(pvals_helminth(data, "IL4_single"))
kable(pvals_helminth(data, "IL13_single")) 
kable(pvals_helminth(data, "IL4_IL13"))

```

\newpage

#Lineage "Agnostic" Single and Double Cytokine Positive CD4+ T Cells
I don't really believe that this is an appropriate way to look at the data because the cells are technically shared between some of these groupings, but I'm almost certain someone would ask the question "What if you didn't consider Il-4 and IL-13 in defining Gamma and TNF single and double positive cells" so I did it for them.

```{r TH1, echo=FALSE, fig.cap="Breakdown of TH1 Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH1 plus explanations.png")
```

```{r TH2, echo=FALSE, fig.cap="Breakdown of TH2 Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH2 plus explanations.png")
```

```{r Single and Double Producers Lin Ag graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)

SP_Ag<-select(Cytokine, Donor, TB, SM, Stim, G_plus, T_plus, IL4_plus, IL13_plus, G_T_plus, IL4_IL13_plus)
melt<-melt(SP_Ag,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_plus", "T_plus", "G_T_plus", "IL4_plus", "IL13_plus", "IL4_IL13_plus"))
melt<-filter(melt, !is.na(label))

### PMA
#TH1 plots
data<-filter(melt, Stim=="PMA" & label %in% c("G_plus", "T_plus", "G_T_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PMA" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_Ag, Stim=="PMA")
kable(pvals_helminth(data, "G_plus"), caption = "Non Specific CD4+ T cell Responses")
kable(pvals_helminth(data, "T_plus"))
kable(pvals_helminth(data, "G_T_plus"))
kable(pvals_helminth(data, "IL4_plus"))
kable(pvals_helminth(data, "IL13_plus"))
kable(pvals_helminth(data, "IL4_IL13_plus"))

### WCL
#TH1 plots
data<-filter(melt, Stim=="WCL" & label %in% c("G_plus", "T_plus", "G_T_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="WCL" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_Ag, Stim=="WCL")
kable(pvals_helminth(data, "G_plus"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals_helminth(data, "T_plus"))
kable(pvals_helminth(data, "G_T_plus"))
kable(pvals_helminth(data, "IL4_plus"))
kable(pvals_helminth(data, "IL13_plus"))
kable(pvals_helminth(data, "IL4_IL13_plus"))

### Peptide
#TH1 plots
data<-filter(melt, Stim=="PEP" & label %in% c("G_plus", "T_plus", "G_T_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PEP" & label %in% c("IL4_plus", "IL13_plus", "IL4_IL13_plus"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_Ag, Stim=="PEP")
kable(pvals_helminth(data, "G_plus"), caption = "Mtb Peptide Specific CD4+ T cell Responses")
kable(pvals_helminth(data, "T_plus"))
kable(pvals_helminth(data, "G_T_plus"))
kable(pvals_helminth(data, "IL4_plus"))
kable(pvals_helminth(data, "IL13_plus"))
kable(pvals_helminth(data, "IL4_IL13_plus"))
```