### 2019-05-22 Cohort-Characteristics.Rmd
includes functional script for processing OG data

---
title: "Cohort_Characteristics"
author: "Taryn McLaughlin"
date: "4/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
library(dplyr)
library(ggplot2)
library(knitr)
library(ggpubr)
```

```{r load data}
DonorInfo<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Donor_Info_clean.csv")
DonorInfo<-filter(DonorInfo, SM %in% c("X", "SM"))
DonorInfo$SM<-factor(DonorInfo$SM, levels=c("X", "SM"))
DonorInfo$TB<-factor(DonorInfo$TB, levels=c("HC", "LTBI", "TB"))
DonorInfo$disease<-paste(DonorInfo$TB, DonorInfo$SM, sep="_")
DonorInfo$disease<-factor(DonorInfo$disease, levels=c("HC_X", "HC_SM",
                                                      "LTBI_X", "LTBI_SM",
                                                      "TB_X", "TB_SM"))
my_comparisons <- list(c("HC_SM", "HC_X"), c("LTBI_SM", "LTBI_X"), c("TB_SM", "TB_X"), 
                       c("HC_SM", "LTBI_SM"), c("TB_SM", "LTBI_SM"), c("HC_SM", "TB_SM"),
                       c("HC_X", "LTBI_X"), c("TB_X", "LTBI_X"), c("HC_X", "TB_X"))
DonorInfo$SchistosomaIntensity[DonorInfo$SM=="X"]<-NA
```

Generating stuff by hand
```{r by hand}
IQR_vals <- function(datatable, column) {
    library(data.table)
    y<-split(datatable, datatable$disease)
    x<-lapply(y, function(y) quantile(y[,column], probs=c(0.5, 0.25, 0.75), na.rm=TRUE))
    #quantile is a stats function that takes a vector "x" and returns the value at a specified quantile
    data.frame(t(x))
}
IQR_pvals<-function(data, column){
    x<-split(data,data$TB)
    #does a wilcox test between SM- and SM+ for whatever column you call in the column argument
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"], exact=FALSE, na.action = na.pass))
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value)
    t(pvals)         
}

###N
n<-t(DonorInfo %>% dplyr::count(disease))
names(n)<-n[1,]

## AGE
fit<-lm(age~disease, data=DonorInfo)
total.p<-anova(fit)$`Pr(>F)`
#Age is different
ggplot(DonorInfo, aes(x=disease, y=age)) +
    geom_boxplot(aes(fill=TB, alpha=SM), size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=4)+
    scale_alpha_manual(values=c(0.5,1))+
    scale_fill_manual(values = c("#1a9850" , "#2166ac", "#b2182b"))+
    theme_classic()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons = my_comparisons,  p.adjust="bonferroni", aes(label=..p.adj..))+
    labs(subtitle = "These aren't corrected for whatever reason will figure out later")
AGE<-cbind(IQR_vals(DonorInfo, "age"),
           cbind(IQR_pvals(DonorInfo, "age"), total.p))
AGE<-dplyr::rename(AGE, HC.p = V1, LTBI.p = V2, TB.p = V3)
AGE$parameter<-"Age"

## SEX
female<-(DonorInfo %>% dplyr::group_by(sex) %>% dplyr::count(disease))[1:6,2:3]
male<-(DonorInfo %>% dplyr::group_by(sex) %>% dplyr::count(disease))[7:12,2:3]
df.sex<-merge(female, male, by="disease")
df.sex<-dplyr::rename(df.sex, female=n.x, male=n.y)
df.sex$femfreq = paste("(",
    round(df.sex$female/(df.sex$female+df.sex$male),2)*100, 
    "%", ")", sep="")
df.sex$malefreq = paste("(",
    round(df.sex$male/(df.sex$female+df.sex$male),2)*100,
    "%", ")", sep="")

total.p<-round(chisq.test(df.sex[,2:3])$p.value, 3)
HC.p<-round(chisq.test(df.sex[1:2,2:3])$p.value, 3)
LTBI.p<-round(chisq.test(df.sex[3:4,2:3])$p.value, 3)
TB.p<-round(chisq.test(df.sex[5:6,2:3])$p.value, 3)
pvals<-data.frame(cbind(HC.p, LTBI.p, TB.p, total.p))
female<-paste(df.sex$female, df.sex$femfreq, sep=" ")
male<-paste(df.sex$male, df.sex$malefreq, sep=" ")
d1<-data.frame(rbind(female, male))
names(d1)<-df.sex$disease
SEX<-cbind(d1, pvals)
SEX$parameter<-row.names(SEX)

##RACE
RACE<-paste(n[2,], "(100%)", sep=" ")

## QFT
qft<-filter(DonorInfo, TB!="TB")
qft$TB<-factor(qft$TB, levels=c("HC", "LTBI"))
qft$disease<-factor(qft$disease, levels=c("HC_X","HC_SM",
                                          "LTBI_X", "LTBI_SM"))
fit<-lm(QFT~disease, qft)
total.p<-anova(fit)$`Pr(>F)`
QFT<-IQR_vals(qft, "QFT")
x<-split(qft,qft$TB)
#does a wilcox test between SM- and SM+ for whatever column you call in the column argument
A<-lapply(x, function(g) wilcox.test(g$QFT~g$SM, exact=FALSE, na.action = na.pass))
pvals<-c(A$HC$p.value, A$LTBI$p.value, total.p)

## Schisto
eggs<-filter(DonorInfo, SM=="SM")
fit<-lm(SchistosomaIntensity~TB, eggs)
eggs.p<-anova(fit)$`Pr(>F)`[1]
EGGS<-cbind(IQR_vals(eggs, "SchistosomaIntensity"),eggs.p)

kable(AGE)
kable(SEX)
kable(QFT)
kable(EGGS)
```

https://cran.r-project.org/web/packages/tableone/vignettes/introduction.html

```{r tableone, echo = FALSE}
DonorInfo$SM<-gsub("SM", "SM+", DonorInfo$SM)
DonorInfo$SM<-gsub("X", "SM-", DonorInfo$SM)

library(tableone)
#Create a variable list which we want in Table 1
listVars <- c("age", "sex", "QFT", "SchistosomaIntensity")

#Define categorical variables
catVars <- c("TB","SM")

mediqr<- c("age", "QFT", "SchistosomaIntensity")

#Total Population
table1 <- CreateTableOne(vars = listVars, data = DonorInfo, factorVars = catVars,
                         strata = c("TB", "SM"))
test<- print(table1, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
             printToggle = FALSE)

tableHC <- CreateTableOne(vars = listVars, data = filter(DonorInfo, TB=="HC"), factorVars = catVars, strata = "SM")
testHC<- print(tableHC, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
             printToggle = FALSE)
tableLTBI <- CreateTableOne(vars = listVars, data = filter(DonorInfo, TB=="LTBI"), factorVars = catVars, strata = "SM")
testLTBI<- print(tableLTBI, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces =
                     TRUE,printToggle = FALSE)
tableTB <- CreateTableOne(vars = listVars, data = filter(DonorInfo, TB=="TB"), factorVars = catVars, strata = "SM")
testTB<- print(tableTB, showAllLevels = TRUE, nonnormal = mediqr, quote = FALSE, noSpaces = TRUE,
             printToggle = FALSE)

write.csv(test, file = "myTable.csv")
kable(test, caption = "FULL")
kable(testHC, caption = "Healthy Controls")
kable(testLTBI, caption = "LTBI")
kable(testTB, caption = "TB")
```

http://api.rpubs.com/isaac_subirana/compareGroups-demo

```{r compare groups, echo = FALSE}

library(compareGroups)
res.total<-descrTable(~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo)
res <- compareGroups(SM ~ age + sex + QFT + SchistosomaIntensity, data = DonorInfo, 
                     method = c(age = 2, QFT = 2, SchistosomaIntensity = 2, sex = 3))

restab <- createTable(res, type = 1, hide.no = "no", show.n = FALSE)
test<-strataTable(restab, "TB")
export2md(test, header.labels = c(p.overall = "p-value"))
export2md(res.total)

```

Plotting stuff against age and sex
```{r age and sex vs, cache = TRUE}
compass<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_Compass_scores_complete.csv")
CD4<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/CD4_clean.csv")
CD4_cyt<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/clean/CD4_Cytokine_clean.csv")

test<-merge(CD4, DonorInfo, by="Donor")
test<-filter(test, Stim=="UN")%>%
    select(-Donor, -X.x, -Sample, -TB.x, -TB.y, -SM.x, -SM.y, -Stim, -X.y, -Excluded)
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column)
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}
### By Cytokine
test<-merge(CD4_cyt, DonorInfo, by="Donor")
test<-filter(test, Stim=="WCL")%>%
    select(-disease, -Excluded, -Cells.Vial, -Viability, -SM.y,- TB.y, -X.y, -SM.x,- TB.x, -X.x,-Sample)
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="WCL")
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}

test<-merge(CD4_cyt, DonorInfo, by="Donor")
test<-filter(test, Stim=="PEP")%>%
    select(-disease, -Excluded, -Cells.Vial, -Viability, -SM.y,- TB.y, -X.y, -SM.x,- TB.x, -X.x,-Sample)
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="PEP")
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}

## COMPASS
test<-merge(compass, DonorInfo, by="Donor")
test<-filter(test, Stim=="WCL", celltype=="CD4")
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="WCL")
    g
}
test.age.plot(test, "FS")
test.age.plot(test, "PFS")

test<-merge(CD4_cyt, DonorInfo, by="Donor")
test<-filter(test, Stim=="PEP")
test.age.plot<-function(datatable, column){
    g<-ggplot(datatable, aes(x=age, y=datatable[,column], color=sex))
    g<- g + geom_point()+theme_bw()+geom_smooth(method="lm", show.legend=TRUE)
    g<- g + labs(title=column, subtitle="PEP")
    g
}
listnames<-names(test)
for(l in listnames){
    print(l)
    p<-test.age.plot(test, l)
    print(p)
}
```