### 2019-05-22 OG-Analysis.Rmd
basic Mtb CD4 analysis

---
title: "OG-Analysis"
author: "Taryn McLaughlin"
date: "5/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in data files}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
CD4_summary<-read.csv("OG_CD4-bulk_summary.csv")
CD4_summary$SM<-factor(CD4_summary$SM, levels=c("X", "SM"))
CD4_OG_summary<-read.csv("OG_CD4-OG_summary.csv")
CD4_OG_summary$SM<-factor(CD4_OG_summary$SM, levels=c("X", "SM"))
PBMC<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/exported/OG_PBMC.csv")
```

```{r functions and stuff}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/Plotting_1902.R")
```

```{r PBMC}
PBMC<-filter(PBMC, !is.na(Study.ID))
PBMC<-filter(PBMC, Helminth.Status %in% c("Negative", "Schisto+"))
PBMC$Helminth.Status<-factor(PBMC$Helminth.Status, levels=c("Negative", "Schisto+"))
PBMC$TB.Status<-factor(PBMC$TB.Status, levels=c("Healthy Control", "LTBI", "Active TB"))
PBMC<-filter(PBMC, !(Study.ID=="NK2208" & Exp.Name=="TM181216"))
PBMC<-filter(PBMC, !(Study.ID=="NK2306" & Exp.Name=="TM181206"))
PBMC<-filter(PBMC, !(Study.ID=="NK2353" & Exp.Name=="TM190411"))
PBMC<-filter(PBMC, !(Study.ID=="NK2384" & Exp.Name=="TM190411"))
PBMC$TB<-PBMC$TB.Status
PBMC$SM<-PBMC$Helminth.Status

grid.arrange(plot_6(PBMC, "SM", "Viability")+labs(y="Viability"),
    plot_6(PBMC, "SM", "Cells.Vial")+labs(y="Cell Count"),
    ncol=2
)

plot_pairs<-function(data, xval, yval){
    ggplot(data, aes(x=data[,xval], y=data[,yval]))+
        geom_point(size=3, aes(color=TB, alpha=SM)) +
        geom_line(aes(group=Study.ID))+
        scale_color_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
        scale_alpha_manual(values=c(0.5,1))+
        facet_grid(~SM)+
        theme_bw()+ theme(legend.position="bottom")+
        theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
        theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
        facet_grid(SM~TB, scale="free")+
        stat_compare_means(label = "p.format", p.adjust.method = "fdr", hide.ns = TRUE, size = 4, 
                           method="wilcox.test", paired = TRUE, label.y= (max(data[,yval])*1.1))
}

melt<-melt(PBMC, id.vars=c("Study.ID", "TB", "SM"), measure.vars = c("Viability", "OG.Viability"))
plot_pairs(melt, "variable", "value")+labs(y="Viability", x="Time")

melt<-melt(PBMC, id.vars=c("Study.ID", "TB", "SM"), measure.vars = c("Cells.Vial", "OG.Cell.."))
plot_pairs(melt, "variable", "value")+labs(y="Count", x="Time")
```

```{r}
library(dplyr)
antigen.ref<-MIMOSA.Results
antigen.ref$id<-rownames(antigen.ref)
antigen.ref$id<-gsub("_Treatment", "", antigen.ref$id)
antigen.ref<-arrange(antigen.ref, id)

phenotype<-CD4_OG_summary
phenotype$id<-paste(phenotype$Donor, phenotype$Stim, sep="_")
phenotype<-arrange(phenotype, id)

remove<-subset(antigen.ref, antigen.ref[,"Pr.response"]<.7)$id
print(length(remove)) #see how many sample ids will be excluded
numeric.only <- function(X,...){
      returnCols <- names(X)
      a<-sapply(X, is.numeric)
      print(returnCols[a == "TRUE"])
}

phenotype[(phenotype$id %in% remove), numeric.only(phenotype)]<-NA
```

#Pep
```{r}
df<-filter(CD4_summary, Stim=="PEP")
top<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+
    theme(text = element_text(size=20)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    stat_compare_means(label = "p.format")+
    coord_cartesian(ylim=c(2,30))

bottom<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB)) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(strip.background = element_blank(),strip.text.x = element_blank())+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    coord_cartesian(ylim=c(0,2))

grid.arrange(top, bottom, ncol=1, heights=c(4,8))

df<-filter(phenotype, Stim=="PEP", TB!="HC")
test<-filter(phenotype, Stim=="PEP", TB!="HC", !is.na(IFNg_freq))
kable(xtabs(~TB+SM, test))
df$SM<-factor(df$SM, levels=c("X", "SM"))
grid.arrange(plot_2(df, "SM", "IFNg_freq")+labs(y="IFNg+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "TNFa_freq")+labs(y="TNFa+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "IL4_freq")+labs(y="IL-4+ Proliferating CD4+ T cells (%)", x=""),
             ncol=3

)

grid.arrange(plot_2(df, "SM", "CXCR3_freq")+labs(y="CXCR3+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "CCR4_freq")+labs(y="CCR4+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "Tbet_freq")+labs(y="Tbet+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "GATA3_freq")+labs(y="GATA3+ Proliferating CD4+ T cells (%)", x=""),
             ncol=4

)
```

#WCL
```{r}
df<-filter(CD4_summary, Stim=="WCL")
top<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), 
        outlier.shape = NA, aes(fill=TB)) +geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank())+
    theme(text = element_text(size=20)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    stat_compare_means(label = "p.format")+
    coord_cartesian(ylim=c(5,50))

bottom<-ggplot(df, aes(x=SM, y=OG_freq, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), 
                 outlier.shape = NA, aes(fill=TB)) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_fill_manual(values =c("#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,1))+
    facet_grid(~TB)+
    theme_classic()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    theme(strip.background = element_blank(),strip.text.x = element_blank())+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB)+
    labs(y="", x="")+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    coord_cartesian(ylim=c(0,5))

grid.arrange(top, bottom, ncol=1, heights=c(4,6))

df<-filter(phenotype, Stim=="WCL", TB!="HC")
test<-filter(phenotype, Stim=="WCL", TB!="HC", !is.na(IFNg_freq))
kable(xtabs(~TB+SM, test))

grid.arrange(plot_2(df, "SM", "IFNg_freq")+labs(y="IFNg+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "TNFa_freq")+labs(y="TNFa+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "IL4_freq")+labs(y="IL-4+ Proliferating CD4+ T cells (%)", x=""),
             ncol=3

)

grid.arrange(plot_2(df, "SM", "CXCR3_freq")+labs(y="CXCR3+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "CCR4_freq")+labs(y="CCR4+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "Tbet_freq")+labs(y="Tbet+ Proliferating CD4+ T cells (%)", x=""),
             plot_2(df, "SM", "GATA3_freq")+labs(y="GATA3+ Proliferating CD4+ T cells (%)", x=""),
             ncol=4

)
```