### 2019-05-29 PBMC.Rmd
markdown for donor info and PBMC info

---
title: "PBMC_Donor"
author: "Taryn McLaughlin"
date: "5/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
```

```{r}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
```

```{r viability and cell counts}
library(ggplot2)
Donor$TB[grepl("HD", Donor$StudyID)==TRUE]<-"Yerkes"
palette.four<-c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b")
grid.arrange(ggplot(Donor, aes(x=SM, y=Viability, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="Viability", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Live Cells (%)"),

ggplot(Donor, aes(x=SM, y=Cells.Vial, fill=TB, alpha=SM)) +
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape=NA) +
    geom_jitter(width=.1,height=0, shape=16,size=3)+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_fill_manual(values = c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    theme_bw()+ theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(angle=90, vjust=0.6)) +   
    facet_grid(~TB, scales="free")+
    labs(title="Cell Counts", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Cells/mL)")
)
```

```{r}
#This was to pick the most robust responders for something else so Idk
library(dplyr)
mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
HD<-filter(mimosa, grepl("HD", mimosa$Sample))%>%
    filter(Stim!="PMA")
HDscores<-aggregate(score~Donor, HD, FUN=sum)
HDkeep<-head(arrange(HDscores, score), 10)
#382 479 487 are BCG
HDremove<-HD$Donor[!(HD$Donor %in% HDkeep$Donor)]
CD4<-dplyr::filter(CD4, !(Donor %in% HDremove))
MFI<-dplyr::filter(MFI, !(Donor %in% HDremove))
Compass<-dplyr::filter(Compass, !(Donor %in% HDremove))
Donor<-dplyr::filter(Donor, !(Donor %in% HDremove))
```