### 2019-06-06 PBMC.Rmd
Including Oregon Green PBMC data

---
title: "PBMC_Donor"
author: "Taryn McLaughlin"
date: "5/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)
library(data.table)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(dplyr)
```

```{r source scripts}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
```

```{r ICS prep}
library(dplyr)
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
#Import information about the samples from my lab notes
ics<-read.csv("ICS_PBMC.csv")
#remove samples that are other helminths since we dropped that group
#Not US, US and Unknown are from the Yerkes donors
    ics<-dplyr::filter(ics, Helminth.Status=="Schisto+" | Helminth.Status== "Negative" | Helminth.Status== "Not US" |  Helminth.Status=="US" | Helminth.Status== "Unknown")
#take out all the stuff that isn't consistent like the "TM" in experiment name and the % in viability

    ics$Viability <- gsub("%/%" , "", ics$Viability)
    ics$Viability<-as.numeric(gsub("%" , "" , ics$Viability))
    ics$Exp.Date <- gsub("TM", "", ics$Exp.Date)

#Remove samples that are being excluded from the ICS analysis
ics<- dplyr::filter(ics, !(Excluded=="EX"))%>%
    dplyr::rename(TB = TB.Status, SM = Helminth.Status, Donor = Study.ID, Exp.Name=Exp.Date)%>% 
    dplyr::select(Donor, TB, SM, Exp.Name, Viability, Cells.Vial, Excluded) #take only the columns you need

ics$Donor<-tstrsplit(ics$Donor, "-")[[1]]

ics$TB<-as.character(ics$TB)
for(i in 1:length(ics$TB)){
    if (ics$TB[i]=="Active TB") ics$TB[i]<-"TB"
    else if (ics$TB[i]=="Healthy Control") ics$TB[i]<-"HC"
    else if (ics$TB[i]=="LTBI") ics$TB[i]<-"LTBI"
    else ics$TB[i]<-"N"
}

ics$SM<-as.character(ics$SM)
for(i in 1:length(ics$SM)){
    if (ics$SM[i]=="Negative") ics$SM[i]<-"X"
    else if (ics$SM[i]=="Schisto+") ics$SM[i]<-"SM"
    else ics$SM[i]<-"N"
}

write.csv(ics, "/Applications/Old Computer/Day Lab/Flow-Data/clean/ICS_DonorInfo_clean.csv")

```

```{r OG prep}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported")
og<-read.csv("OG_PBMC.csv")
og<-rename(og, Donor=Study.ID, TB=TB.Status, SM=Helminth.Status)%>%
    filter(Donor!="", Excluded!="EX")%>%
    select(Donor, TB, SM, Exp.Name, Viability, Cells.Vial, OG.Viability, OG.Cell.., Excluded, Exp.Name, Cheryl)

og$Donor<-tstrsplit(og$Donor, "-")[[1]]
og$Donor<-gsub("TRU", "", og$Donor)

og$TB<-as.character(og$TB)
for(i in 1:length(og$TB)){
    if (og$TB[i]=="Active TB") og$TB[i]<-"TB"
    else if (og$TB[i]=="Healthy Control") og$TB[i]<-"HC"
    else if (og$TB[i]=="LTBI") og$TB[i]<-"LTBI"
    else og$TB[i]<-"N"
}

og$SM<-as.character(og$SM)
for(i in 1:length(og$SM)){
    if (og$SM[i]=="Negative") og$SM[i]<-"X"
    else if (og$SM[i]=="Schisto+") og$SM[i]<-"SM"
    else og$SM[i]<-"N"
}

write.csv(og, "/Applications/Old Computer/Day Lab/Flow-Data/clean/OG_DonorInfo_clean.csv")
```

```{r combine}
pbmc<-merge(ics, og, by=c("Donor", "TB", "SM"), all=TRUE)
names(pbmc)<-c("Donor", "TB" , "SM" , 
             "ICS_Exp.Name" , "ICS_Viability", "ICS_Cells.Vial", "ICS_Exluded",
             "OG_Exp.Name" , "PreOG_Viability" , "PreOG_Cells.Vial" , "PostOG_Viability" , "PostOG_Cells.Vial" , "OG_Excluded" , "Cheryl") 
```

```{r Import Kenya data}
#Import the clinical enrollment data from the cleaned merged epi file
    QFT<- read.csv("/Applications/Old Computer/Epi Project/Data_clean/complete-kenya-filledin.csv")
#The studyIDs of coenrolled people are NK#_TRU# and we want just the NK# to match the lab notes
    QFT$Donor<-tstrsplit(QFT$StudyID, "_")[[1]]
#trim the data set to only those study IDs in the lab notes and select the columns you want
    QFT<-dplyr::select(QFT, Donor, age, sex, QFT, SchistosomaIntensity)%>%
        dplyr::filter(Donor %in% pbmc$Donor)
```

```{r merge all}

#Merge the two tables together and keep all observations since the yerkes donors won't be in the epi data
    DonorInfo<-merge(pbmc, QFT, by="Donor", all = TRUE)
#Manual Edits that may get deleted after Kenya script is finalized
    DonorInfo$TB[DonorInfo$Donor=="NK2183"]<-as.character("HC")
    DonorInfo$QFT[DonorInfo$Donor=="NK2323"]<-0.04
    DonorInfo$SM[DonorInfo$Donor=="NK2325"]<-as.character("Unknown")

#Used in
    DonorInfo$used<-NA
    for(i in 1:length(DonorInfo$used)){
        if(!is.na(DonorInfo$ICS_Exp.Name[i]) & !is.na(DonorInfo$OG_Exp.Name[i]))
            DonorInfo$used[i]<-"ICS/OG"
        else if (!is.na(DonorInfo$ICS_Exp.Name[i]) & is.na(DonorInfo$OG_Exp.Name[i]))
            DonorInfo$used[i]<-"ICS only"
        else if (is.na(DonorInfo$ICS_Exp.Name[i]) & !is.na(DonorInfo$OG_Exp.Name[i]))
            DonorInfo$used[i]<-"OG only"
        else DonorInfo$used[i]<-NA
    }
write.csv(DonorInfo, "/Applications/Old Computer/Day Lab/Flow-Data/clean/DonorInfo_clean.csv")

```

```{r viability and cell counts, fig.height=5, fig.width=10}
pbmc$SM<-gsub("SM", "SM+", pbmc$SM)
pbmc$SM<-gsub("X", "SM-", pbmc$SM)
pbmc$SM<-factor(pbmc$SM, levels=c("N", "SM-", "SM+"))
pbmc$TB<-factor(pbmc$TB, levels=c("N", "HC", "LTBI", "TB"))

grid.arrange(
    plot_pbmc(pbmc, "ICS_Viability")+labs(y="Viability"),
    plot_pbmc(pbmc, "ICS_Cells.Vial")+labs(y="Cells/Vial"),
    top=text_grob("ICS PBMCs", size=20), ncol=2
)

grid.arrange(
    plot_pbmc(pbmc, "PreOG_Viability")+labs(y="Viability"),
    plot_pbmc(pbmc, "PreOG_Cells.Vial")+labs(y="Cells/Vial"),
    top=text_grob("Pre OG Staining", size=20), ncol=2
)

grid.arrange(
    plot_pbmc(pbmc, "PreOG_Viability")+labs(y="Viability"),
    plot_pbmc(pbmc, "PreOG_Cells.Vial")+labs(y="Cells/Vial"),
    top=text_grob("Post OG Staining", size=20), ncol=2
)
```

```{r}
melted<-melt(pbmc, id.vars = c("Donor", "TB", "SM"), 
             measure.vars = c("PreOG_Viability" , "PostOG_Viability"))
plot_pbmc.pairs(melted)
melted<-melt(pbmc, id.vars = c("Donor", "TB", "SM"), 
             measure.vars = c("PreOG_Cells.Vial", "PostOG_Cells.Vial"))
plot_pbmc.pairs(melted)
```

```{r, eval=FALSE}
#This was to pick the most robust responders for something else so Idk
library(dplyr)
mimosa<-read.csv("/Applications/Old Computer/Day Lab/Flow-Data/Clean/CD4_MIMOSA_Complete.csv")
HD<-filter(mimosa, grepl("HD", mimosa$Sample))%>%
    filter(Stim!="PMA")
HDscores<-aggregate(score~Donor, HD, FUN=sum)
HDkeep<-head(arrange(HDscores, score), 10)
#382 479 487 are BCG
HDremove<-HD$Donor[!(HD$Donor %in% HDkeep$Donor)]
CD4<-dplyr::filter(CD4, !(Donor %in% HDremove))
MFI<-dplyr::filter(MFI, !(Donor %in% HDremove))
Compass<-dplyr::filter(Compass, !(Donor %in% HDremove))
Donor<-dplyr::filter(Donor, !(Donor %in% HDremove))
```