### 2019-07-02 ICS_HC.Rmd
Includes boolean graphs for phenotype

---
title: "ICS_HC"
author: "Taryn McLaughlin"
date: "6/2/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
```

```{r source}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
cd4<-plot_filter(read.csv("ICS_CD4_clean.csv"))

CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))

phen<-plot_filter(read.csv("ICS_CD4_phenotyping_clean_NAs.csv"))
```

##Actual analysis

```{r fig.width=16, fig.height=8}
HC<-mutate_hc(subset(cd4, cd4$Stim=="UN"))
test<-melt(HC, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = grep("._._._.", names(cd4)))
test2<-filter(test, variable!="x_x_x_x")
ggplot(test2, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot()+
    scale_alpha_manual(values=c(0.5,1))+
    scale_fill_manual(values =c("#1a9850"))+
    stat_compare_means(label="p.signif", hide.ns = TRUE)+
    theme(legend.position="none")+
    labs(title="Lineage Marker Expression on Total CD4+ T cells in Healthy Adults")
```

```{r Figure 1a, }
HC<-mutate_hc(subset(cd4, cd4$Stim=="UN"))
HC_T<-ggplot(HC, aes(x=SM, y=Tbet))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$Tbet)*1.1),
                       label.x.npc = "center")+
    labs(title="",
         x="",
         y=paste("Total T-bet+ CD4+ T cells (%)"))

HC_T3<-ggplot(HC, aes(x=SM, y=TbetCXCR3))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +   
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TbetCXCR3)*1.1),
                       label.x.npc = "center")+
    labs(title="",
         x="",
         y=paste("T-bet+CXCR3+ CD4+ T cells (%)"))

HC_T4<-ggplot(HC, aes(x=SM, y=TbetCCR4))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +   
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TbetCXCR3)*1.1),
                       label.x.npc = "center")+
    labs(title="",
         x="",
         y=paste("T-bet+CXCR3+ CD4+ T cells (%)"))

HC_lin<-grid.arrange(HC_T, HC_T3, HC_T4, ncol=3, top=text_grob("Expression of TH1 Lineage Markers", size=20))
HC_lin
```

###Difference aesthetics Same Cytokine Graph
```{r}
HC<-mutate_hc(filter(CD4, Stim=="PMA"))
test<-melt(HC, id.vars=c("Donor", "TB", "SM"), measure.vars = c("TH1.Freq", "TH2.Freq", "TH1.2.Freq"))

one<-ggplot(test, aes(x=SM, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
      geom_point(position=position_jitterdodge(dodge.width=0.9)) +
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE,
                       label.x.npc = "center")+
    facet_wrap(~variable, scales="free")+
    labs(title="Cytokine Frequencies",
         x="",
         y=paste("Freq of Cytokine+ CD4+ T cells (%)"))

two<-ggplot(test, aes(x=SM, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
      geom_point(position=position_jitterdodge(dodge.width=0.9)) +
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE,
                       label.x.npc = "center")+
    facet_wrap(~variable)+
    labs(title="Cytokine Frequencies",
         x="",
         y=paste("Freq of Cytokine+ CD4+ T cells (%)"))

three<-ggplot(test, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
      geom_point(position=position_jitterdodge(dodge.width=0.9)) +
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE,
                       label.x.npc = "center")+
    labs(title="Cytokine Frequencies",
         x="",
         y=paste("Freq of Cytokine+ CD4+ T cells (%)"))

one
two
three
```

##Different Ways to Visualize TH1 and TH2 cell marker expression
```{r Figure 1b, hide=TRUE}
HC<-mutate_hc(filter(phen, Stim=="PMA"));
HC_1T<-ggplot(HC, aes(x=SM, y=TH1_Tbet))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_Tbet)*1.1),
                       label.x.npc = "center")+
    labs(title="TH1 Cells",
         x="",
         y=paste("Total T-bet+ CD4+ T cells (%)"))
HC_2T<-ggplot(HC, aes(x=SM, y=TH2_Tbet))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_Tbet)*1.1),
                       label.x.npc = "center")+
    labs(title="TH2 Cells",
         x="",
         y=paste(""))

HC_1G<-ggplot(HC, aes(x=SM, y=TH1_GATA3))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_GATA3)*1.1),
                       label.x.npc = "center")+
    labs(title="TH1 Cells",
         x="",
         y=paste("Total GATA3+ CD4+ T cells (%)"))
HC_2G<-ggplot(HC, aes(x=SM, y=TH2_GATA3))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_GATA3)*1.1),
                       label.x.npc = "center")+
    labs(title="TH2 Cells",
         x="",
         y=paste(""))

HC_13<-ggplot(HC, aes(x=SM, y=TH1_CXCR3))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_CXCR3)*1.1),
                       label.x.npc = "center")+
    labs(title="TH1 Cells",
         x="",
         y=paste("Total CXCR3+ CD4+ T cells (%)"))
HC_23<-ggplot(HC, aes(x=SM, y=TH2_CXCR3))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH1_CXCR3)*1.1),
                       label.x.npc = "center")+
    labs(title="TH2 Cells",
         x="",
         y=paste(""))

HC_14<-ggplot(HC, aes(x=SM, y=TH1_CCR4))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_CCR4)*1.1),
                       label.x.npc = "center")+
    labs(title="TH1 Cells",
         x="",
         y=paste("Total CXCR3+ CD4+ T cells (%)"))
HC_24<-ggplot(HC, aes(x=SM, y=TH2_CCR4))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA, aes(fill=TB, alpha=SM)) +
    geom_jitter(width=.1,height=0, shape=16,size=2)+
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    scale_x_discrete(labels=c("SM-", "SM+"))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(vjust=0.5)) + 
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 12))+
    stat_compare_means(label = "p.format", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE, label.y= (max(HC$TH2_CCR4)*1.1),
                       label.x.npc = "center")+
    labs(title="TH2 Cells",
         x="",
         y=paste(""))
```

```{r, fig.width=16, fig.height=8}
one.b<-grid.arrange(
    HC_T<-grid.arrange(HC_1T, HC_2T, ncol=2, top=text_grob("T-bet", size=20)),
    HC_3<-grid.arrange(HC_13, HC_23, ncol=2, top=text_grob("CXCR3", size=20)),
    HC_G<-grid.arrange(HC_1G, HC_2G, ncol=2, top=text_grob("GATA3", size=20)),
    HC_4<-grid.arrange(HC_14, HC_24, ncol=2, top=text_grob("CCR4", size=20)),
    ncol=4)
```

```{r}
HC<-mutate_hc(filter(phen, Stim=="PMA"));
vars <- names(HC)[grep("TH1_|TH2_", names(HC))]
vars1 <- c("TH1_Tbet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_Tbet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4")
vars2 <- vars[grep(".P|.N", vars)]
vars3 <- vars[grep("._._._.", vars)]

```

```{r}
test<-melt(HC, id.vars=c("Donor", "TB", "SM"), vars1)
test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine + CD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)

plot_markers<-function(df, marker){
    ggplot(df[df$variable==marker,], aes(x=SM, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
    geom_point(position=position_jitterdodge(dodge.width=0.9)) +
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(text = element_text(size=14)) +
    theme(strip.text = element_text(size = 14))+
    facet_grid(variable~subset, scales="free")+
    labs(y="", x="")+
    stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE,
                       label.x.npc = "center")}

grid.arrange(
    plot_markers(test, "Tbet")+theme(axis.text.x = element_blank()),
    plot_markers(test, "GATA3")
        +theme(axis.text.x = element_blank())+theme(strip.text.x = element_blank()),
    plot_markers(test, "CXCR3")
        +theme(axis.text.x = element_blank())+theme(strip.text.x = element_blank()),
    plot_markers(test, "CCR4")
        +theme(strip.text.x = element_blank())+        scale_x_discrete(labels=c("SM-", "SM+")),
    ncol=1
)
```

```{r fig.width=16, fig.height=10}
test<-melt(HC, id.vars=c("Donor", "TB", "SM"), vars2)
test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine Producing CD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test<-filter(test, variable!="CCR4.N_CXCR3.N" , variable!="GATA3.N_Tbet.N")

ggplot(test, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
    geom_point(position=position_jitterdodge(dodge.width=0.9)) +
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 20))+
    facet_wrap(~subset, ncol=1)+
    stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE,
                       label.x.npc = "center")

test<-melt(HC, id.vars=c("Donor", "TB", "SM"), vars3)
test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine Producing CD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test<-filter(test, -x_x_x_x)

ggplot(test, aes(x=variable, y=value, fill=TB, alpha=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 0.9), outlier.shape = NA) +
    geom_point(position=position_jitterdodge(dodge.width=0.9)) +
    scale_fill_manual(values="#1a9850")+
    scale_alpha_manual(values=c(0.5,1))+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(text = element_text(size=14)) +
    theme(strip.text.x = element_text(size = 20))+
    facet_wrap(~subset, ncol=1)+
    stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 4, 
                       method="wilcox.test", paired = FALSE,
                       label.x.npc = "center")
```

#Proof that we can combine hc and ltbi
##Lineage Markers on UN

```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(cd4, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(cd4, Stim=="UN")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

##Cytokine Analysis from PMA

```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(CD4, Stim=="PMA")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```

#Phenotype from PMA
```{r, fig.height=3, fig.width=6}
library(knitr)
colnames<-names(dplyr::select(phen, -X.1, -X, -id, -Sample, -Donor, -Stim, -TB, -SM))
df<-filter(phen, Stim=="PMA")
df<-mutate_hc(df)
for(col in colnames){
    print(col)
    print(pvals_hc(df, col))
    grid.arrange(
       plot_hc(df[df$QFT=="QFT-",], "SM", col)+labs(title="QFT- Only"), 
       plot_hc(df, "SM", col)+labs(title="All Healthy Controls"),
       ncol=2
    )
}
```