### 2019-07-29 OG_cleaning.Rmd

---
title: "OG_cleaning"
author: "Taryn McLaughlin"
date: "6/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/exported/")
#find all the csv filenames in that directory and put them in a list
filenames =list.files(pattern="OG")
#for each name in that list, assign it to the output of read.csv of that name
#aka read in each file by it's name
for (i in 1:length(filenames)) 
    assign(filenames[i], read.csv(filenames[i]))
```

```{r source functions used in this script}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/OG-cleaning-functions.R")
```

```{r}
l<-lapply(ls(pattern="summary.csv"), function(x) get(x))

l<-lapply(l, function(x) split_og(x))
l<-lapply(l, function(x) clean_ognames(x))
l<-lapply(l, function(x) remove_samples(x))
l<-lapply(l, function(x) remove_donors(x))
l<-lapply(l, function(x) changeTB(x))

filenames<-ls(pattern="summary.csv")
for(i in 1:15){
assign(filenames[i], data.frame(l[i]))}
```

```{r create new combo variables from exported boolean phenotype}
l<-lapply(ls(pattern="bulk_summary"), function(x) get(x))
l<-lapply(l, function(x) sum_boolean(x))
filenames<-ls(pattern="bulk_summary")
for(i in 1:5){
assign(filenames[i], data.frame(l[i]))}

l<-lapply(ls(pattern="_OG"), function(x) get(x))
l<-lapply(l, function(x) sum_boolean(x))
filenames<-ls(pattern="_OG")
for(i in 1:9){
assign(filenames[i], data.frame(l[i]))}
```

```{r}
l<-lapply(ls(pattern="bulk_summary"), function(x) get(x))
l<-lapply(l, function(x) OG_background.subtract(x, "Donor", "Stim", "UN", "OG_freq"))
filenames<-ls(pattern="bulk_summary")
for(i in 1:5){
assign(filenames[i], data.frame(l[i]))}
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")

write.csv(OG_CD3_bulk_summary.csv, "OG_CD3_bulk_summary_clean.csv")
write.csv(OG_CD3_OG_summary.csv, "OG_CD3_OG_summary_clean.csv")

write.csv(OG_CD4_bulk_summary.csv, "OG_CD4_bulk_summary_clean.csv")
write.csv(OG_CD4_OG_summary.csv, "OG_CD4_OG_summary_clean.csv")
write.csv(OG_CD4_OG.GT_summary.csv, "OG_CD4_OG.GT_summary_clean.csv")
write.csv(OG_CD4_OG.IL4_summary.csv, "OG_CD4_OG.IL4_summary_clean.csv")
write.csv(OG_CD4_OG.TH1_summary.csv, "OG_CD4_OG.TH1_summary_clean.csv")
write.csv(OG_CD4_OG.TNFa_summary.csv, "OG_CD4_OG.TNFa_summary_clean.csv")
write.csv(OG_CD4_OG.IFNg_summary.csv, "OG_CD4_OG.IFNg_summary_clean.csv")

write.csv(OG_CD8_bulk_summary.csv, "OG_CD8_bulk_summary_clean.csv")
write.csv(OG_CD8_OG_summary.csv, "OG_CD8_OG_summary_clean.csv")

write.csv(OG_gd_bulk_summary.csv, "OG_gd_bulk_summary_clean.csv")
write.csv(OG_GDonly_bulk_summary.csv, "OG_GDonly_bulk_summary_clean.csv")
write.csv(OG_gd_OG_summary.csv, "OG_gd_OG_summary_clean.csv")
write.csv(OG_GDonly_OG_summary.csv, "OG_GDonly_OG_summary_clean.csv")

write.csv(OG_counts_summary.csv, "OG_counts_summary_clean.csv")
```