### 2019-09-12 PaperFigures.Rmd
added in space for flow plots and letters for labeling

---
title: "Paper Figures"
author: "Taryn McLaughlin"
date: "8/28/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
```

```{r plot functions}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")

baseplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ theme(legend.position="none")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x="", y="")
}

booleanplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(text = element_text(size=16), axis.text.x = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
    labs(x="", y="")
}

allplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
    geom_boxplot(size=1, position=position_dodge2(width=0.2, padding = 0.2, preserve = "single"), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none") +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    facet_wrap(~variable)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5))+
    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10)+
    labs(x="", y="")
}

cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))

all_comps<-list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-"))

healthy.cols<- c("SM+" = "#1a9850", "SM-" ="#91cf60")
ltbi.cols<- c("SM+" = "#2166ac", "SM-" = "#85ABD1")
tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
all.cols<- c("Naive" = "#C8D0D8","HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
             "LTBI SM+" = "#2166ac", "LTBI SM-" = "#85ABD1",
             "TB SM+" = "#b2182b", "TB SM-"="#E59EA6")
```

```{r load data}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
CD4<-read.csv("ICS_CD4_clean.csv")
CYT<-read.csv("ICS_CD4_Cytokine_clean.csv")
PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
OG<-data.table(read.csv("OG_CD4_bulk_summary_clean.csv"))
OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
OG.IFNg<-read.csv("OG_CD4_OG.IFNg_summary_clean_NAs.csv")
OG.GT<-read.csv("OG_CD4_OG.GT_summary_clean_NAs.csv")
OG.th1<-read.csv("OG_CD4_OG.TH1_summary_clean_NAs.csv")
luminex<-read.csv("/Applications/Old Computer/Day Lab/Luminex/Clean Data/Luminex_combined_background.txt")
```

#All groups
Statistics were performed using the Mann-Whitney U test for non-parametric data. P-values <0.05 were considered significant.
```{r}
cd4<-plot_filter(subset(CD4, CD4$Stim=="UN"))
cyt<-plot_filter(subset(CYT, CYT$Stim=="PMA"))
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
```

#Figure 1
```{r, fig.width=15, fig.height=8}
melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))

tf<-allplot(melt, "TB", "value", all.cols)+    
    labs(y="(%) Total CD4+ T cells")

melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))

sm<-allplot(melt, "TB", "value", all.cols) +    
    labs(y="(%) Total CD4+ T cells")

first_row = cowplot::plot_grid(NULL, NULL, labels = c('A' , 'B'), nrow=1, label_size = 20, rel_widths = c(3,2))
second_row = cowplot::plot_grid(tf, sm,labels = c('C', 'D'), nrow=1, label_size = 20)
figure1<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
```

```{r pvals for ifng and tnfa}
kable(table(cd4$disease), caption="Total CD4 Phenotypes")
kable(table(cyt$disease), caption="Total CD4 Cytokine")

y<-split(cyt,cyt$SM) 
A<-lapply(y, function(g) kruskal.test(IFNg~TB, g))
B<-lapply(y, function(g) kruskal.test(TNFa~TB, g))

pval<-c(A$`SM-`$p.value, A$`SM+`$p.value, B$`SM-`$p.value, B$`SM+`$p.value)
comparison<-c("SM- IFNγ", "SM+ IFNγ", "SM- TNFα", "SM+ TNFα")
kable(cbind(comparison, pval), caption="Stats across Tb groups")
```

#Figure 2
####Generic cytokine production from CD4+ T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
```{r, fig.width=15, fig.height=8}
melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))

one<-allplot(melt, "TB", "value", all.cols)+ 
    facet_wrap(~variable, labeller = cyt_groups)+
    labs(y="(%) Total CD4+ T cells")

melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))

two<-allplot(melt, "TB", "value", all.cols)+
    facet_wrap(~variable, labeller = cyt_groups)+
    coord_cartesian(ylim=c(0,8))+
    labs(y="(%) Total CD4+ T cells")

melt<-melt(subset(CYT, CYT$Stim=="PMA"), 
           id.vars=c("TB","SM"), 
           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))

cyt<-filter_hc(melt)
a<-baseplot(cyt, "SM", "value", healthy.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    labs(y="(%) Total CD4+ T cells")+coord_cartesian(ylim=c(0,45))+
    labs(title="Healthy Controls")

cyt<-filter_ltbi(melt)
b<-baseplot(cyt, "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,45))+
    labs(title="LTBI")

cyt<-filter_tb(melt)
c<-baseplot(cyt, "SM", "value", tb.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,45))+
    labs(title="TB")

first_row = cowplot::plot_grid(NULL, NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(one, two, labels = c('C', 'D'), nrow=1, label_size = 20)
third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('E'), label_size = 20)
figure2 = cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
```

#Figure 3
####Generic TH1 and TH2 cytokine producing cells are stratified by lineage-specific transcription factors, and CCR4
Amongst TH1 cytokine producing CD4+ T cells, moderate frequencies express the transcription factor T-bet while very low frequencies express GATA3. In contrast, amongst TH2 cytokine producing CD4+ T cells, low frequencies express T-bet while high frequencies express GATA3. Similarly high frequencies of TH2 cytokine producing cells express CCR4 compared to TH1 cytokine producing cells. Both TH1 and TH2 cells have equivalent moderate frequencies of CXCR3 expressing cells. There are no differences by S. mansoni status.

```{r, fig.width=15, fig.height=8}
phen<-filter_hc(subset(PHEN, PHEN$Stim=="PMA"))
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

a<-booleanplot(test, "SM","value", healthy.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="(%) Cytokine+ CD4+ T cells",
         title="HC")

phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

b<-booleanplot(test, "SM","value", ltbi.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="(%) Cytokine+ CD4+ T cells",
         title="LTBI")

phen<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+\nCD4+ T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

c<-booleanplot(test, "SM","value", tb.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    labs(y="(%) Cytokine+ CD4+ T cells",
         title="TB")

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('B', 'C', 'D'), label_size = 20)
figure3 <- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,4))
```

#LTBI
#Figure 4
####LTBI individuals coinfected with S. mansoni have higher Mtb peptide-specific cytokine production, specifically due to higher production of IFNγ
```{r, narrow down ltbi data}
cyt<-filter_ltbi(subset(CYT, CYT$Stim=="PEP"))
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))

kable(table(cyt$disease), caption="Pep cytokine")
kable(table(phen$disease[!is.na(phen$TH1_MFI_CCR4)]), caption="Pep phenotype")
```

```{r cytokine, fig.width=15, fig.height=8}
melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
a<-baseplot(melt, "SM", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = cyt_groups)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.15))+
    coord_cartesian(ylim=c(0,.15))+
    labs(y="(%) CD4+ T cells")

melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
b<-baseplot(melt, "SM", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = cyt_groups)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y = (.95*.02))+
    coord_cartesian(ylim=c(0,.02))

```

####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status
```{r ics main four markers, fig.width=15, fig.height=8}
melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("ifng_T-bet", "ifng_GATA3", "ifng_CXCR3", "ifng_CCR4"))
melt$variable<-gsub("ifng_", "", melt$variable)
melt$variable <- factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*70))+
    coord_cartesian(ylim=c(0,70))+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")

sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
    coord_cartesian(ylim=c(0,100))+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(a, b, nrow=1, labels=c('B', 'C'), label_size = 20)
third_row = cowplot::plot_grid(tf, sm, nrow=1, labels=c('D', 'E'), label_size = 20)
figure4 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
```

#TB 
#Figure 5
```{r, narrow down tb data}
cyt<-filter_tb(subset(CYT, CYT$Stim=="PEP"))
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))

kable(table(cyt$disease), caption="Pep cytokine")
kable(table(phen$disease[!is.na(phen$TH1_MFI_CCR4)]), caption="Pep phenotype")
```

####Similar to QFT+, individuals with active TB, co-infected with S. mansoni maintain function Mtb-specific CD4+ T cell cytokine production. There are higher frequencies of TH1/2 cytokine co-producing CD4+ T cells amongst SM+ individuals, however the frequency of these cells is incredibly low. The dominant cytokine subgroup is still TH1 cytokine producing cells.
*Again this is where I need an aesthetics opinion*

```{r, fig.width=15, fig.height=8}
melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
a<-baseplot(melt, "disease", "value", tb.cols)+    
    facet_wrap(~variable, labeller = cyt_groups)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.15))+
    coord_cartesian(ylim=c(0,.15))+
    labs(y="(%) CD4+ T cells")

melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
b<-baseplot(melt, "disease", "value", tb.cols)+ 
    facet_wrap(~variable, labeller = cyt_groups)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.95*.02))+
    coord_cartesian(ylim=c(0,.02))
```

####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers. There are numerous phenotypic differences amongst Mtb-specific TH1 cytokine producing cells betweeen SM- and SM+ groups. While T-bet is still the dominant transcription factor expressed in both groups, most differences are due to higher frequencies of GATA3 and CCR4 co-expressing cells.

```{r main four markers, fig.width=15, fig.height=8}
melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4"))
melt$variable<-gsub("TH1_", "", melt$variable)
melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", tb.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
    coord_cartesian(ylim=c(0,80))+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")

sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", tb.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*80))+
    coord_cartesian(ylim=c(0,80))

phen$TH1_GATA4notCCR4<-rowSums(phen[,grep("TH1_._g_._x", names(phen))])
melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_GATA3CCR4", "TH1_GATA4notCCR4"))

groups <- as_labeller(c("TH1_GATA3CCR4" = "GATA3+CCR4+", 
                    "TH1_GATA4notCCR4" = "GATA3+CCR4-"))

both<-baseplot(melt, "SM", "value", tb.cols)+    
    facet_wrap(~variable, labeller = groups)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3)+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")

tbcyt<-cowplot::plot_grid(NULL, a, b, nrow=1, labels=c("A","B", "C"), label_size = 20, rel_widths = c(1.5,1,1))
tbphen<-cowplot::plot_grid(tf, sm, both, nrow=1, labels=c("D","E","F"), label_size = 20)
figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,3))
```

#Proliferation Assay
#Figure 6
##All Groups Responses to SEB
```{r}
og<-plot_filter(dplyr::filter(OG, Stim=="SEB"))
ogphen<-plot_filter(dplyr::filter(OGPHEN, Stim=="SEB"))

kable(table(og$disease), caption="Total Proliferation")
kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="SEB phenotype")
```

```{r, proliferation and total cyt, fig.width=15, fig.height=8}
groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4"))

part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
og<-allplot(part1, "TB", "value", all.cols)+
    facet_wrap(~variable, scales="free", labeller = groups)+
    labs(y="(%) Total CD4+ T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))

one<-allplot(melt, "TB", "value", all.cols)+    
    facet_wrap(~variable, labeller = groups, ncol=2)+
    labs(y="(%) Proliferating CD4+ T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))

two<-allplot(melt, "TB", "value", all.cols)+    
    facet_wrap(~variable, labeller = groups)
```

```{r, marker on proliferating, fig.width=15, fig.height=8}
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))

tf<-allplot(melt, "TB", "value", all.cols)+    
    labs(y="(%) Proliferating CD4+ T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))

sm<-allplot(melt, "TB", "value", all.cols)

first_row = cowplot::plot_grid(NULL, labels = c('A', 'B'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
figure6<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
```

#Figure 7
##Mtb-Specific Responses in LTBI
```{r}
og<-filter_ltbi(dplyr::filter(OG, Stim=="PEP"))
ogphen<-filter_ltbi(dplyr::filter(OGPHEN, Stim=="PEP"))
oggt<-filter_ltbi(dplyr::filter(OG.GT, Stim=="PEP"))
ogth1<-filter_ltbi(dplyr::filter(OG.th1, Stim=="PEP"))

kable(table(og$disease), caption="PEP Proliferation")
kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="PEP phenotype")
```

####There are no differences in Mtb-specific proliferation betweeen SM+ and SM- QF+ individuals. The cytokine production from expanded Mtb-specific CD4+ T cells only differs with regards to IFNγ IL-4 co-production, however the frequency of these cells is incredibly low. The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFα. 
```{r og and cyt, fig.width=15, fig.height=8}
groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4"))

melt<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
og<-baseplot(melt, "TB", "value", ltbi.cols)+
    facet_wrap(~variable, scales="free", labeller = groups)+
    coord_cartesian(ylim=c(0,3))+
    labs(y="(%) Total CD4+ T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
one<-baseplot(melt, "TB", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = groups, ncol=2)+
    labs(y="(%) Proliferating CD4+ T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
two<-baseplot(melt, "TB", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = groups)
```

####Similar to the overnight assay, expanded Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status.

*I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
###On proliferating cells
```{r og main four markers, fig.width=16, fig.height=8}
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
    coord_cartesian(ylim=c(0,100))+
    labs(y="(%) Proliferating Mtb-Specific cells")

sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
    coord_cartesian(ylim=c(0,100))

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(og, one, two, nrow=1, rel_widths = c(1.2,2,1), labels=c("B", "C","D"), label_size = 20)
third_row = cowplot::plot_grid(tf, sm, labels=c("E", "F"), nrow=1, label_size = 20)
figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
```

###On TH1 proliferating cells
```{r og th1 main four markers, fig.width=16, fig.height=8}
melt<-melt(ogth1, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

a<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
    coord_cartesian(ylim=c(0,100))+
    labs(y="(%) Proliferating Mtb-Specific cells")

b<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(method="wilcox.test" , label="p.format", size=4, label.x=1.3, label.y=(.98*100))+
    coord_cartesian(ylim=c(0,100))

ogth1<-cowplot::plot_grid(a,b,ncol=2)
```

#Luminex Data

####Furthermore, IFNγ and TNFa are also the two dominant cytokines produced when analyzed by luminex. The concentration of all cytokines measured by luminex are equivalent between SM- and SM+ groups.

*Honestly I will have to just remake the whole thing a couple times and stitch them together to fix the axes so just ignore them for now*

```{r luminex, fig.width=12, fig.height=32}
lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
names(lum)<-gsub("IL", "IL-", names(lum))
names(lum)<-gsub("IFNg", "IFNγ", names(lum))
names(lum)<-gsub("TNFa", "TNFα", names(lum))

melt<-melt(lum, id.vars=c("TB", "SM"), measure.vars=grep("pg", names(lum)))
melt$variable<-gsub("_pg.ml", "", melt$variable)
melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-17A", "IL-22", "IL-10", "IL-4", "IL-5", "IL-13"))

a<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
    labs(y="pg/mL")

b<-baseplot(melt, "SM", "value", ltbi.cols)+facet_wrap(~variable, scales="free")+
    labs(y="pg/mL")+
    coord_cartesian(ylim=c(0,40))

lumplot<-cowplot::plot_grid(a,b, NULL, ncol=1, align = "v", labels=c("A","A", "B"), label_size = 20)
```

####However the correlation between cytokines differs between groups. Of note, the correlation between IFNγ and TNFa is stronger amongst SM+ individuals than SM- individuals.
```{r cor plots}
df<-lum[,3:11]
names(lum)<-gsub("_pg.ml", "", names(lum))
pca1 <- prcomp(df, scale = TRUE)
pcas<-data.frame(pca1$x)
rotations<-data.frame(pca1$rotation)
rotations$cytokine<-row.names(rotations)
melt<-melt(rotations, id.vars="cytokine")

stims<-factor(lum$Stim)
worms<-factor(lum$SM)
a<-ggplot(pcas, aes(x=pcas[,1], y=pcas[,2]))+geom_point(aes(col=worms))+labs(y="PCA2", x="PCA1")

neg <- cor(lum[lum$SM=="SM-",3:11], method="spearman")
order <- corrMatOrder(neg, order = "hclust")
order.neg <- neg[order, order]
res1 <- cor.mtest(lum[lum$SM=="SM-",3:11], conf.level = .95)
pvals <- res1$p[order, order]
corrplot(order.neg, method = "circle", insig = "label_sig", tl.col = "black", 
         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
         type="upper")

pos <- cor(lum[lum$SM=="SM+",3:11], method="spearman")
order.pos <- pos[order, order]
res1 <- cor.mtest(lum[lum$SM=="SM+",3:11], conf.level = .95)
pvals <- res1$p[order, order]
corrplot(order.pos, method = "circle", insig = "label_sig", tl.col = "black", 
         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
         type="upper")
```
#Exporting Data
```{r}
setwd("/Users/tarynam/Desktop/Manuscript1")
cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
cowplot::save_plot("figure2.png", figure2, base_height=14, base_width=16) #Figure 2
cowplot::save_plot("figure3.png", figure3, base_height=14, base_width=16)
cowplot::save_plot("figure4.png", figure4, base_height=14, base_width=16)
cowplot::save_plot("figure5.png", figure5, base_height=14, base_width=16)
cowplot::save_plot("figure6.png", figure6, base_height=16, base_width=16)
cowplot::save_plot("figure7.png", figure7, base_height=16, base_width=16)
cowplot::save_plot("figure8.png", lumplot, base_height=32, base_width=16)
```

##ICS Supplementary
```{r}
booleanplot2<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(text = element_text(size=20), axis.text.x = element_text(size=20))+
    theme(plot.subtitle = element_text(hjust = 0.5)) +
    labs(x="", y="")
}
```

###Boolean PMA
*but I need to make them pretty and I dont' feel like doing that right now*
```{r boolean pma cyt, fig.width=15, fig.height=12}
cyt<-plot_filter(CYT[CYT$Stim=="PMA",])
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")

labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
breaks <- levels(test$variable)

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,3))+
    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
    labs(y="(%) CD4+ T cells")

th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,25))+
    theme(strip.text.y = element_blank(), strip.background.y = element_blank())

th2<-booleanplot2(test[test$th=="TH2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,3))

supp1<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
```

###Boolean Phenotypes
###TH1
```{r pma th1 boolean, fig.width=20, fig.height=18}
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))

melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

df<-data.frame(cbind(levels(test$variable),T,G,c3,c4))
dfm<-melt(df, measure.vars=c("T", "G","c3","c4"))
dfm$variable<-factor(dfm$variable, levels=c("c4","c3","G", "T"))

data_table <- ggplot(dfm, aes(x = V1, y = variable, label = format(value))) +
    geom_text(size = 10) + 
    theme_bw() + 
    labs(x="",y="")+
    theme(strip.text.x = element_blank(), strip.background.x = element_blank())+
    scale_y_discrete(labels=c("T-bet", "GATA3", "CXCR3", "CCR4"))+
    theme(axis.text.y = element_text(size=20))+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks = element_blank())

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp2a<-booleanplot2(test, "variable", "value", all.cols)+
    facet_grid(TB~tf, scales="free_x")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,65))

```
###TH2  
```{r pma th2 boolean, fig.width=20, fig.height=18}
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))

melt<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH2_._._._.", names(phen)))
test<-filter(melt, variable!="TH2_x_x_x_x")
test$variable<-gsub("TH2_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp2b<-booleanplot2(test, "variable", "value", all.cols)+
    facet_grid(TB~tf, scales="free_x")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,65))
```

```{r}
supp2<-cowplot::plot_grid(supp2a, supp2b, nrow=2, labels=c("A", "B"))
```

###Peptide-Specific Cytokine Production  
```{r, fig.width=15, fig.height=8}
melt<-melt(subset(CYT, CYT$Stim=="PEP"), 
           id.vars=c("TB","SM"), 
           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))

cyt<-filter_ltbi(melt)
b<-baseplot(cyt, "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,.2))

cyt<-filter_tb(melt)
c<-baseplot(cyt, "SM", "value", tb.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,.2))

supp3<-cowplot::plot_grid(b, c, ncol=3)
```

```{r, fig.width=16, fig.height=8}
cyt<-plot_filter(CYT[CYT$Stim=="PEP",])
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

G   <-c("IFNγ +", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
il4 <-c("IL-4 +", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
il13<-c("IL-13 +", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
T   <-c("TNFα +", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")

labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
breaks <- levels(test$variable)

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

supp4a<-booleanplot2(test[test$TB=="LTBI",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
    coord_cartesian(ylim=c(0,.16))+
    labs(y="(%) CD4+ T cells")

supp5a<-booleanplot2(test[test$TB=="TB",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
    coord_cartesian(ylim=c(0,.15))+
    labs(y="(%) CD4+ T cells")
```

###Peptide Mtb-specific TH1 Cytokine+ Cell Phenotypes
```{r ics phenotyping, fig.width=16, fig.height=8}
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))

melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp4b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")
```

```{r, fig.width=15, fig.height=8}
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))

#graph phenotype
melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp5b<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    labs(y="(%) Mtb-Specific TH1 Cytokine+ cells")+
    coord_cartesian(ylim=c(0,55))+
    stat_compare_means(method="wilcox.test",label="p.signif", hide.ns=TRUE, label.y=50, size=10)
```

###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
```{r, fig.width=12, fig.height=6}
total<-filter_ltbi(subset(CD4, CD4$Stim=="UN"))
total$cell<-"Total"
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
mtb_th1<-select(phen, names)
names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
mtb_th1$cell<-"Mtb-Specific"
long<-rbindlist(list(total, mtb_th1), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
```{r, fig.width=12, fig.height=6}
pma<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
pma$cell<-"PMA"
pep<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
pep$cell<-"Mtb"
long<-rbindlist(list(pma, pep), fill=TRUE)
names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
long<-select(long, names)
names(long)<-gsub("TH1_", "", names(long))

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

###Comparing Mtb-specific TH1 cells to total CD4 T cells in peptide stim  
```{r, fig.width=12, fig.height=6}
total<-filter_tb(subset(CD4, CD4$Stim=="UN"))
total$cell<-"Total"
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
mtb_th1<-select(phen, names)
names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
mtb_th1$cell<-"Mtb-Specific"
long<-rbindlist(list(total, mtb_th1), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = tb.cols)+
    scale_x_discrete(labels=c("Total", "Mtb", "Total", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")

```

###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
```{r, fig.width=12, fig.height=6}
pma<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
pma$cell<-"PMA"
pep<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
pep$cell<-"Mtb"
long<-rbindlist(list(pma, pep), fill=TRUE)
names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
long<-select(long, names)
names(long)<-gsub("TH1_", "", names(long))

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))

ggplot(melt, aes(x=x, y=value, fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    facet_wrap(~variable, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = tb.cols)+
    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

```{r}
supp4<-cowplot::plot_grid(supp4a, supp4b, nrow=2, labels=c("A", "B"))
supp5<-cowplot::plot_grid(supp5a, supp5b, nrow=2, labels=c("A", "B"))
```

##Proliferation Supplementary
```{r}
ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
test<-melt[-grep("._._._.", melt$variable),]
test<-filter(test, variable!="x_x_x")
test$variable<-factor(test$variable, levels=c(
"T_G_x" ,       
"T_x_x" ,         
"x_G_x" ,         
"T_G_4" ,
"T_x_4" ,         
"x_G_4" ,         
"x_x_4"          
))
test$group<-"TH1/2"
for(i in 1:length(test$group)){
    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
        test$group[i]<- "TH1"
    }
    else if (test$variable[i] == "x_x_4"){
        test$group[i]<- "TH2"
    }
}

T <- c("+", "+", "-", "+", "+", "-", "-")
G <- c("+", "-", "+", "+", "-", "+", "-")
il4 <- c("-", "-", "-", "+", "+", "+", "+")

labels <- paste0(T, "\n", G, "\n", il4)
breaks <- levels(test$variable)

supp6a<-booleanplot(test, "variable", "value", ltbi.cols)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    facet_grid(~group, scales="free_x", space="free_x")+
    labs(y="(%) Proliferating CD4+ T cells")

```

```{r og phenotype, fig.width=16, fig.height=8}
#graph phenotype
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
test<-filter(melt, variable!="x_x_x_x")
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" ,       
"t_x_3_x" ,         
"t_x_x_4" ,         
"t_x_x_x" ,
"t_g_3_4" ,         
"t_g_3_x" ,         
"t_g_x_4" ,         
"t_g_x_x" ,
"x_g_3_4" ,         
"x_g_3_x" ,         
"x_g_x_4" ,         
"x_g_x_x" ,
"x_x_3_4" ,          
"x_x_3_x" ,         
"x_x_x_4"           
))

T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
breaks <- levels(test$variable)

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp6b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=6)+
    labs(y="(%) Proliferating Mtb-Specific cells")
```

```{r}
supp6<-cowplot::plot_grid(supp6a, supp6b, nrow=2, labels=c("A", "B"), rel_widths = c(3,4))
```

###Comparing Mtb-specific TH1 cells to PMA TH1 cells  
```{r, fig.width=12, fig.height=6}
un<-filter_ltbi(subset(OG, OG$Stim=="UN"))
un$cell<-"total"
seb<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="SEB"))
seb$cell<-"seb"
long<-rbindlist(list(un, seb), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))

ggplot(long, aes(x=cell, y=long[,"T-bet"], color=disease))+
    geom_point(size=3, alpha=0.5)+geom_line(aes(group=Donor))+
    facet_wrap(~TB, scales="free", nrow=1)+
    theme_classic()+
    scale_fill_manual(values = ltbi.cols)+
    scale_x_discrete(labels=c("PMA", "Mtb", "PMA", "Mtb"))+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=16))+
    theme(plot.title = element_text(hjust = 0.5)) +
    stat_compare_means(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")) 
                       ,method="wilcox.test", label="p.format")+
    labs(x="", y="")
```

```{r}
lum_x<-lum[lum$SM=="SM-",]
df_x<-as.matrix(lum_x[,3:11])
row.names(df_x)<-lum_x$Sample
hc_x <- hclust(as.dist(1-cor(df_x)))
x<-plot(hc_x)

lum_sm<-lum[lum$SM=="SM+",]
df_sm<-as.matrix(lum_sm[,3:11])
row.names(df_sm)<-lum_sm$Sample
hc_sm <- hclust(as.dist(1-cor(df_sm)))
sm<-plot(hc_sm)

df<-as.matrix(lum[,3:11])
#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
```

#Exporting Data
```{r}
setwd("/Users/tarynam/Desktop/Manuscript1")
cowplot::save_plot("supp1.png", supp1, base_height=14, base_width=16)
cowplot::save_plot("supp2.png", supp2, base_height=28, base_width=16)
cowplot::save_plot("supp3.png", supp3, base_height=7, base_width=16)
cowplot::save_plot("supp4.png", supp4, base_height=14, base_width=16)
cowplot::save_plot("supp5.png", supp5, base_height=14, base_width=16)
cowplot::save_plot("supp6.png", supp6, base_height=14, base_width=16)
```