### 2019-09-15 gamma-delta.Rmd
pretty figures

---
title: "gamma-delta-figures"
author: "Taryn McLaughlin"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
PBMC<-dplyr::select(read.csv("DonorInfo_clean.csv"), -X)
    PBMC<-PBMC[grep("OG", PBMC$used),]
    plot_3(PBMC[PBMC$SM=="SM",], "SchistosomaIntensity")+
    labs(y="Eggs/gram")
```

#ICS Assay

```{r, fig.width=12, fig.height=8}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
compass<-read.csv("ICS_compass_complete_clean.csv")
compass$TB<-factor(compass$TB, levels=c("N", "HC", "LTBI", "TB"))
kable(pvals_SM(compass[compass$Stim=="SWAP",], "FS"), caption="SWAP")
kable(pvals_SM(compass[compass$Stim=="SEA",], "FS"), caption="SEA")

include<-subset(compass, compass$Stim %in% c("SEA", "SWAP"))
plot_cell(include, "FS")+
    facet_grid(Stim~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    labs(y="Functionality Score",
         title="Includes SM-")

exclude<-subset(include, include$SM!="X")
keep<-plot_cell(exclude, "FS")+
    facet_grid(Stim~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "CD4-CD8-"))+
    labs(y="Functionality Score",
         title="Excludes SM-")
```

#Proliferation Assay
```{r load og data}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean")
gd<-dplyr::select(read.csv("OG_gd_bulk_summary_clean.csv"), -X, -X.1)
gd_OG<-dplyr::select(read.csv("OG_gd_OG_summary_clean_NAs.csv"), -X)
GD<-dplyr::select(read.csv("OG_GDonly_bulk_summary_clean.csv"), -X, -X.1)
GD_OG<-dplyr::select(read.csv("OG_GDonly_OG_summary_clean_NAs.csv"), -X)
CD3<-dplyr::select(read.csv("OG_CD3_bulk_summary_clean.csv"), -X, -X.1)
CD3_OG<-dplyr::select(read.csv("OG_CD3_OG_summary_clean.csv"), -X, -X.1)
CD4<-dplyr::select(read.csv("OG_CD4_bulk_summary_clean.csv"), -X, -X.1)
CD8<-dplyr::select(read.csv("OG_CD8_bulk_summary_clean.csv"), -X, -X.1)

mimosa<-dplyr::select(read.csv("OG_MIMOSA_clean.csv"), -X)
    mimosa<-dplyr::filter(mimosa, celltype=="GD" & Exprate==0.01)
```

*I was conflicted so I ran a test to see if there were statistical differences between sm- and sm+ and when there are I plot both. All the plots on the left will include the SM- donors, whereas the plots on the right will exclude them.  I still have to export the proliferation data for the yerkes donors but that's a good chunk of time/effort so I'm waiting until the other paper is further along*

###Starting Frequency of Gamma Delta Cells
```{r}
include<-subset(CD3, CD3$Stim=="UN")
kable(pvals_SM(include, "GD"))
a<-plot_3(include, "GD")+
    coord_cartesian(ylim=c(0,35))+
    labs(y="(%) CD3+ T cells",
         title="Includes SM-")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "GD")+
    coord_cartesian(ylim=c(0,35))+
    labs(y="(%) CD3+ T cells",
         title="Excludes SM-")

cowplot::plot_grid(a,b)
```

###Proportion of proliferating Gamma Delta Cells
```{r, fig.width=12, fig.height=8}
include<-subset(CD3_OG, CD3_OG$Stim=="SWAP")
colnames<-c("CD4", "CD8", "GD")
for(col in colnames){
    y<-pvals_SM(include, col)
    if(dim(y)[1]>0){
        print(col)
        print(y)}
}

include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
melt<-melt(include, id.vars=c("TB", "SM", "Stim"), measure.vars=colnames)
melt$variable<-gsub("GD", "γδ", melt$variable)

melt$celltype<-gsub("", "",melt$variable)

a<-plot_3(melt, "value")+
    facet_grid(Stim~variable)+
    #coord_cartesian(ylim=c(0,65))+
    labs(y="(%) Proliferating T cells",
         title="Includes SM-")

exclude<-subset(melt, melt$SM=="SM")
b<-plot_3(melt, "value")+
    facet_grid(Stim~variable, labeller = as_labeller(c("CD4"="CD4", "CD8"="CD8", "GD"="γδ")))+
    labs(y="(%) Proliferating T cells",
         title="Excludes SM-")

a
```

```{r, fig.width=12, fig.height=8}
include<-subset(CD3_OG, CD3_OG$Stim%in%c("SEA","SWAP"))
melt<-melt(include, id.vars=c("TB", "SM", "Stim"), measure.vars=c("CD4", "CD8", "GD"))
melt$celltype<-gsub("", "",melt$variable)

a<-plot_cell(melt, "value")+
    facet_grid(Stim~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
    labs(y="(%) Proliferating+ T cells",
         title="Includes SM-")

exclude<-subset(melt, melt$SM=="SM")
b<-plot_cell(melt, "value")+
    facet_grid(Stim~TB)+
    scale_x_discrete(labels = c("CD4", "CD8", "γδ"))+
    labs(y="(%) Proliferating+ T cells",
         title="Excludes SM-")

a
```

###Proliferation of Gamma Delta Cells
```{r}
include<-subset(gd, gd$Stim=="SEA")
pvals_SM(include, "OG")
a<-plot_3(include, "OG")+
    facet_grid(~Stim)+
    labs(y="(%) γδ T cells")

include<-subset(gd, gd$Stim=="SWAP")
pvals_SM(include, "OG")
b<-plot_3(include, "OG")+
    facet_grid(~Stim)

cowplot::plot_grid(a,b)
```

###Proliferation of CD4 T Cells
```{r}
include<-subset(CD4, CD4$Stim=="SEA")
pvals_SM(include, "OG")
a<-plot_3(include, "OG")+
    facet_grid(~Stim)+
    labs(y="(%) CD4 T cells")

include<-subset(CD4, CD4$Stim=="SWAP")
pvals_SM(include, "OG")
b<-plot_3(include, "OG")+
    facet_grid(~Stim)+
    labs(y="")

cowplot::plot_grid(a,b)
```

###Proliferation of CD8 T Cells
```{r}
include<-subset(CD8, CD8$Stim=="SEA")
pvals_SM(include, "OG")
a<-plot_3(include, "OG")+
    facet_grid(~Stim)+
    labs(y="(%) CD8 T cells")

include<-subset(CD8, CD8$Stim=="SWAP")
pvals_SM(include, "OG")
b<-plot_3(include, "OG")+
    facet_grid(~Stim)+
    labs(y="")

cowplot::plot_grid(a,b)
```

###IL-4
```{r}
include<-subset(gd_OG, gd_OG$Stim=="SWAP")
kable(pvals_SM(include, "IL4"))
a<-plot_3(include, "IL4")+
    labs(y="(%) Proliferating γδ T cells")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "IL4")+
    labs(y="(%) Proliferating γδ T cells")+
    labs(y="")

cowplot::plot_grid(a,b)
```

###TNFa
```{r}
include<-subset(gd_OG, gd_OG$Stim=="SWAP")
kable(pvals_SM(include, "TNFa"))
a<-plot_3(include, "TNFa")+
    labs(y="(%) Proliferating γδ T cells")

exclude<-subset(include, include$SM=="SM")
b<-plot_3(exclude, "TNFa")+
    labs(y="(%) Proliferating γδ T cells")

cowplot::plot_grid(a,NULL)
```

###All Cytokine Combinations
```{r}
test<-subset(gd_OG, gd_OG$Stim=="SWAP")
vars<-names(test)[grep("._._.", names(test))]
vars<-vars[-grep("._._._.", vars)]
for(v in vars){
    y<-pvals_SM(test, v)
    if(dim(y)[1]>0){
        print(v)
        print(y)}
}

include<-melt(test, id.vars=c("Donor", "TB", "SM", "Stim"), measure.vars = vars)
include<-filter(include, variable!="x_x_x")
include$variable<-factor(include$variable, levels=c(
"T_G_x" ,       
"T_x_x" ,         
"x_G_x" ,         
"T_G_4" ,
"T_x_4" ,         
"x_G_4" ,         
"x_x_4"          
))

T <- c("+", "+", "-", "+", "+", "-", "-")
G <- c("+", "-", "+", "+", "-", "+", "-")
il4 <- c("-", "-", "-", "+", "+", "+", "+")

labels <- paste0(T, "\n", G, "\n", il4)
breaks <- levels(include$variable)

ggplot(include, aes(x=variable, y=value, fill=TB))+
    geom_boxplot(outlier.shape = NA)+
    scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
    scale_x_discrete(labels=labels, breaks=breaks)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Includes SM-",
         y="(%) Proliferating γδ T cells")

exclude<-subset(include, include$SM=="SM")
ggplot(exclude, aes(x=variable, y=value, fill=TB))+
    geom_boxplot(outlier.shape = NA)+
    scale_fill_manual(values =c("#00AFBB", "#E7B800", "#FC4E07"))+
    scale_x_discrete(labels=labels, breaks=breaks)+
    stat_compare_means(label="p.signif", hide.ns = TRUE, size=8)+
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title="Excludes SM-",
         y="(%) Proliferating γδ T cells")
```