### 2019-09-28 PaperFigures.Rmd
separated main figures from supplementary figures

---
title: "Paper Figures"
author: "Taryn McLaughlin"
date: "8/28/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
```

```{r plot functions}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")

baseplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ theme(legend.position="none")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(text = element_text(size=20),
          axis.text.x = element_text(size = 16),
          strip.text = element_text(size = 20))+
    theme(plot.title = element_text(hjust = 0.5, size=20)) +
    labs(x="", y="")
}

booleanplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=SM))+
    geom_boxplot(size=1, position=position_dodge(width = 1), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none")+
    theme(text = element_text(size=20), 
          axis.text.x = element_text(size=16), 
          strip.text = element_text(size = 20))+
    theme(plot.title = element_text(hjust = 0.5, size=20)) +
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
    labs(x="", y="")
}

allplot<-function(data, xval, yval, colors){
    ggplot(data, aes(data[,xval], data[,yval], fill=disease))+
    geom_boxplot(size=1, position=position_dodge2(width=0.2, padding = 0.2, preserve = "single"), outlier.shape = NA)+
    scale_fill_manual(values = colors)+
    theme_classic()+ 
    theme(legend.position="none") +
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    facet_wrap(~variable)+
    theme(text = element_text(size=20),
            axis.text.x = element_text(size = 16),
            strip.text = element_text(size = 20))+
    theme(plot.title = element_text(hjust = 0.5, size=20))+
    labs(x="", y="")
}

cyt_groups <- as_labeller(c("IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4", "IL13" = "IL-13"))

all_comps<-list(c("HC SM+", "HC SM-"),c("LTBI SM+", "LTBI SM-"), c("TB SM+", "TB SM-"))

healthy.cols<- c("SM+" = "#1a9850", "SM-" ="#91cf60")
ltbi.cols<- c("SM+" = "#2166ac", "SM-" = "#85ABD1")
tb.cols<- c("SM+" = "#b2182b", "SM-"="#E59EA6")
all.cols<- c("Naive" = "#C8D0D8","HC SM+" = "#1a9850", "HC SM-" ="#91cf60",
             "LTBI SM+" = "#2166ac", "LTBI SM-" = "#85ABD1",
             "TB SM+" = "#b2182b", "TB SM-"="#E59EA6")
```

```{r load data}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
CD4<-read.csv("ICS_CD4_clean.csv")
CYT<-read.csv("ICS_CD4_Cytokine_clean.csv")
PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
OG<-data.table(read.csv("OG_CD4_bulk_summary_clean.csv"))
OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
OGnope<-read.csv("OG_CD4_noOG_summary_clean.csv")
OG.IFNg<-read.csv("OG_CD4_OG.IFNg_summary_clean_NAs.csv")
OG.GT<-read.csv("OG_CD4_OG.GT_summary_clean_NAs.csv")
OG.th1<-read.csv("OG_CD4_OG.TH1_summary_clean_NAs.csv")
luminex<-read.csv("/Applications/Old Computer/Day Lab/Luminex/Clean Data/Luminex_combined_background.txt")
```

#All groups
Statistics were performed using the Mann-Whitney U test for non-parametric data. P-values <0.05 were considered significant.
```{r}
cd4<-plot_filter(subset(CD4, CD4$Stim=="UN"))
cyt<-plot_filter(subset(CYT, CYT$Stim=="PMA"))
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))
```

#Figure 1
```{r, fig.width=15, fig.height=8}
melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))

tf<-allplot(melt, "TB", "value", all.cols)+    
    labs(y="(%) Total CD4 T cells")+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10, label.y=25)+
    coord_cartesian(ylim=c(0,30))

melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))

sm<-allplot(melt, "TB", "value", all.cols) +    
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="")+
    stat_compare_means(method = 'wilcox.test', label="p.signif", hide.ns=TRUE, size=10, label.y=60)

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,2))
second_row = cowplot::plot_grid(tf, sm,labels = c('B', ''), nrow=1, label_size = 20)
figure1<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
figure1
```

#Figure 1 alt
```{r, fig.width=15, fig.height=8}
melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))

tf<-allplot(melt, "disease", "value", all.cols)+    
    labs(y="(%) Total CD4 T cells")+
    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", 
                              "LTBI\nSM-", "LTBI\nSM+", 
                              "TB\nSM-", "TB\nSM+"))+
    stat_compare_means(comparisons=list(c("HC SM-", "HC SM+")), method = 'wilcox.test', label="p.format", size=6, label.y=.9*30)+
    coord_cartesian(ylim=c(0,30))

melt<-melt(cd4, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))

sm<-allplot(melt, "disease", "value", all.cols) +    
    labs(y="")+    
    scale_x_discrete(labels=c("HC\nSM-", "HC\nSM+", 
                              "LTBI\nSM-", "LTBI\nSM+", 
                              "TB\nSM-", "TB\nSM+"))+
    stat_compare_means(comparisons=list(c("TB SM-", "TB SM+")), 
                       method = 'wilcox.test', label="p.format", size=6, label.y=.9*65)+
    coord_cartesian(ylim=c(0,65))

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,2))
second_row = cowplot::plot_grid(tf, sm,labels = c('B', ''), nrow=1, label_size = 20)
figure1alt<- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1,2))
figure1alt
```

```{r pvals for ifng and tnfa}
kable(table(cd4$disease), caption="Total CD4 Phenotypes")
kable(table(cyt$disease), caption="Total CD4 Cytokine")

y<-split(cyt,cyt$SM) 
A<-lapply(y, function(g) kruskal.test(IFNg~TB, g))
B<-lapply(y, function(g) kruskal.test(TNFa~TB, g))

pval<-c(A$`SM-`$p.value, A$`SM+`$p.value, B$`SM-`$p.value, B$`SM+`$p.value)
comparison<-c("SM- IFNγ", "SM+ IFNγ", "SM- TNFα", "SM+ TNFα")
kable(cbind(comparison, pval), caption="Stats across Tb groups")
```

#Figure 2
####Generic cytokine production from CD4 T cells stimulated with PMA & Ionomycin does not differ between SM- and SM+ Healthy Adults
```{r, fig.width=15, fig.height=15}
melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))

one<-allplot(melt, "TB", "value", all.cols)+ 
    facet_wrap(~variable, labeller = cyt_groups)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="(%) Total CD4 T cells")

melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))

two<-allplot(melt, "TB", "value", all.cols)+
    facet_wrap(~variable, labeller = cyt_groups)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    coord_cartesian(ylim=c(0,8))+
    labs(y="")

melt<-melt(subset(CYT, CYT$Stim=="PMA"), 
           id.vars=c("TB","SM"), 
           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))

cyt<-filter_hc(melt)
a<-baseplot(cyt, "SM", "value", healthy.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    labs(y="(%) Total CD4 T cells")+coord_cartesian(ylim=c(0,45))+
    labs(title="HC")

cyt<-filter_ltbi(melt)
b<-baseplot(cyt, "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,45))+
    labs(title="LTBI")

cyt<-filter_tb(melt)
c<-baseplot(cyt, "SM", "value", tb.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,45))+
    labs(title="TB")

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(one, two, labels = c('B', ''), nrow=1, label_size = 20)
third_row = cowplot::plot_grid(a, b, c, nrow=1, labels=c('C'), label_size = 20)
figure2 = cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(1,2,2))
figure2
```

some stats comparing th1 to th2 for reference. 
```{r}
x<-filter(plot_filter(melt), variable!="TH1.2.Freq")
y<-split(x,list(x$SM, x$TB)) 
A<-lapply(y, function(g) wilcox.test(value~variable, g))

pval<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
        A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
        A$`SM-.TB`$p.value, A$`SM+.TB`$p.value)
group<-c("HC SM-", "HC SM+", "LTBI SM-", "LTBI SM+", "TB SM-", "TB SM+")
table<-data.frame(cbind(group, pval))
table$pval<-as.numeric(as.character(table$pval))
table$pval[table$pval<.0001]<-"<0.0001"
kable(table, caption="TH1 to TH2")
```

some stats comparing th1 to th1/2 for reference. 
```{r}
x<-filter(plot_filter(melt), variable!="TH2.Freq")
y<-split(x,list(x$SM, x$TB)) 
A<-lapply(y, function(g) wilcox.test(value~variable, g))

pval<-c(A$`SM-.HC`$p.value, A$`SM+.HC`$p.value,
        A$`SM-.LTBI`$p.value, A$`SM+.LTBI`$p.value, 
        A$`SM-.TB`$p.value, A$`SM+.TB`$p.value)
group<-c("HC SM-", "HC SM+", "LTBI SM-", "LTBI SM+", "TB SM-", "TB SM+")
table<-data.frame(cbind(group, pval))
table$pval<-as.numeric(as.character(table$pval))
table$pval[table$pval<.0001]<-"<0.0001"
kable(table, caption="TH1 to TH1/2")
```

#Figure 3
####Generic TH1 and TH2 cytokine producing cells are stratified by lineage-specific transcription factors, and CCR4
Amongst TH1 cytokine producing CD4 T cells, moderate frequencies express the transcription factor T-bet while very low frequencies express GATA3. In contrast, amongst TH2 cytokine producing CD4 T cells, low frequencies express T-bet while high frequencies express GATA3. Similarly high frequencies of TH2 cytokine producing cells express CCR4 compared to TH1 cytokine producing cells. Both TH1 and TH2 cells have equivalent moderate frequencies of CXCR3 expressing cells. There are no differences by S. mansoni status.

```{r, fig.width=15, fig.height=8}
phen<-filter_hc(subset(PHEN, PHEN$Stim=="PMA"))
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+\nCD4 T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

a<-booleanplot(test, "SM","value", healthy.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+

    labs(y="(%) Cytokine+ CD4 T cells",
         title="HC")

phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+\nCD4 T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

b<-booleanplot(test, "SM","value", ltbi.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(plot.margin=margin(l=-1,unit="cm"))+
    labs(y="",
         title="LTBI")

phen<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
test<-melt(phen, id.vars=c("TB", "SM"), 
           measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4",
           "TH2_T-bet", "TH2_GATA3", "TH2_CXCR3", "TH2_CCR4"))

test$subset<-tstrsplit(test$variable, "_")[[1]]
test$subset<-paste(test$subset, "Cytokine+\nCD4 T Cells", sep=" ")
test$variable<-gsub("TH1_","", test$variable)
test$variable<-gsub("TH2_","", test$variable)
test$variable<-factor(test$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

c<-booleanplot(test, "SM","value", tb.cols)+
    facet_grid(variable~subset, scales="free")+
    theme(panel.border = element_rect(fill = NA, colour = "black"))+
    theme(plot.margin=margin(l=-1,unit="cm"))+
    labs(y="",
         title="TB")

second_row <- cowplot::plot_grid(a,b,c, nrow=1)
figure3 <- cowplot::plot_grid(NULL, second_row, labels = c('A', 'B'), label_size = 20, ncol=1, rel_heights = c(1.5,4.5))
figure3
```

#LTBI
#Figure 4
####LTBI individuals coinfected with S. mansoni have higher Mtb peptide-specific cytokine production, specifically due to higher production of IFNγ
```{r, narrow down ltbi data}
cyt<-filter_ltbi(subset(CYT, CYT$Stim=="PEP"))
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))

kable(table(cyt$disease), caption="Pep cytokine")
kable(table(phen$disease[!is.na(phen$TH1_MFI_CCR4)]), caption="Pep phenotype")
```

```{r cytokine, fig.width=15, fig.height=8}
melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
a<-baseplot(melt, "SM", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = cyt_groups)+
    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y = (.90*.15))+
    coord_cartesian(ylim=c(0,.15))+
    labs(y="(%) CD4 T cells")

melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
b<-baseplot(melt, "SM", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = cyt_groups)+
    coord_cartesian(ylim=c(0,.02))

```

####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status
```{r ics main four markers, fig.width=15, fig.height=8}
melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("ifng_T-bet", "ifng_GATA3", "ifng_CXCR3", "ifng_CCR4"))
melt$variable<-gsub("ifng_", "", melt$variable)
melt$variable <- factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,70))+
    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")

sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,100))+
    labs(y="")

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(a, b, tf, sm, nrow=1, labels=c('B', '', 'C', ''), label_size = 20)
figure4opt1 <- cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(3,4))
figure4opt1

#first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
#second_row = cowplot::plot_grid(a, b, nrow=1, labels=c('B', 'C'), label_size = 20)
#third_row = cowplot::plot_grid(tf, sm, nrow=1, labels=c('D', 'E'), label_size = 20)
#figure4opt2 <- cowplot::plot_grid(first_row, second_row, third_row, ncol=1, rel_heights = c(2.5,4,4))

```

#TB 
#Figure 5
```{r, narrow down tb data}
cyt<-filter_tb(subset(CYT, CYT$Stim=="PEP"))
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))

kable(table(cyt$disease), caption="Pep cytokine")
kable(table(phen$disease[!is.na(phen$TH1_MFI_CCR4)]), caption="Pep phenotype")
```

####Similar to QFT+, individuals with active TB, co-infected with S. mansoni maintain function Mtb-specific CD4 T cell cytokine production. There are higher frequencies of TH1/2 cytokine co-producing CD4 T cells amongst SM+ individuals, however the frequency of these cells is incredibly low. The dominant cytokine subgroup is still TH1 cytokine producing cells.
*Again this is where I need an aesthetics opinion*

```{r, fig.width=15, fig.height=8}
melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
a<-baseplot(melt, "SM", "value", tb.cols)+    
    facet_wrap(~variable, labeller = cyt_groups)+
    coord_cartesian(ylim=c(0,.15))+
    labs(y="(%) CD4 T cells")

melt<-melt(cyt, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4", "IL13"))
b<-baseplot(melt, "disease", "value", tb.cols)+ 
    facet_wrap(~variable, labeller = cyt_groups)+
    coord_cartesian(ylim=c(0,.02))
```

####Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers. There are numerous phenotypic differences amongst Mtb-specific TH1 cytokine producing cells betweeen SM- and SM+ groups. While T-bet is still the dominant transcription factor expressed in both groups, most differences are due to higher frequencies of GATA3 and CCR4 co-expressing cells.

```{r main four markers, fig.width=15, fig.height=8}
melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_T-bet", "TH1_GATA3", "TH1_CXCR3", "TH1_CCR4"))
melt$variable<-gsub("TH1_", "", melt$variable)
melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", tb.cols)+    
    facet_wrap(~variable)+
    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=(.98*80))+
    coord_cartesian(ylim=c(0,80))+
    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")

sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", tb.cols)+    
    facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,80))

phen$TH1_GATA4notCCR4<-rowSums(phen[,grep("TH1_._g_._x", names(phen))])
melt<-melt(phen, id.vars=c("TB","SM", "disease"), measure.vars=c("TH1_GATA3CCR4", "TH1_GATA4notCCR4"))

groups <- as_labeller(c("TH1_GATA3CCR4" = "GATA3+CCR4+", 
                    "TH1_GATA4notCCR4" = "GATA3+CCR4-"))

both<-baseplot(melt, "SM", "value", tb.cols)+    
    facet_wrap(~variable, labeller = groups)+
    stat_compare_means(comparisons=list(c("SM-","SM+")), method="wilcox.test" , label="p.format", size=6, label.x=1.3, label.y=.98*13)+
    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")+
    coord_cartesian(ylim=c(0,13))

tbcyt<-cowplot::plot_grid(NULL, a, b, nrow=1, labels=c("A","B", ""), label_size = 20, rel_widths = c(1.5,1,1))
tbphen<-cowplot::plot_grid(tf, sm, both, nrow=1, labels=c("C","","D"), label_size = 20)
figure5<-cowplot::plot_grid(tbcyt, tbphen, nrow=2, rel_heights = c(2,2))
figure5
```

#Proliferation Assay
#Figure 6
##All Groups Responses to SEB
```{r}
groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4"))

og<-plot_filter(dplyr::filter(OG, Stim=="SEB"))
part1<-melt(og, id.vars=c("TB","SM", "disease"), measure.vars="OG")
seb<-allplot(part1, "TB", "value", all.cols)+
    facet_wrap(~variable, scales="free", labeller = groups)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="(%) Total CD4 T cells",
         title="SEB")

og<-plot_filter(dplyr::filter(OG, Stim=="PEP", TB %in% c("LTBI", "TB")))
part2<-melt(og, id.vars=c("TB", "SM", "disease"), measure.vars="OG")
pep<-allplot(part2, "TB", "value", all.cols)+
    facet_wrap(~variable, scales="free", labeller = groups)+
    scale_x_discrete(labels=c("SM- SM+\nLTBI", "SM- SM+\nTB"))+
    coord_cartesian(ylim=c(0,4))+
    labs(y="(%) Total CD4 T cells",
        title="CFP-10 & ESAT-6")

pep2<-allplot(part2, "disease", "value", all.cols)+
    facet_wrap(~variable, scales="free", labeller = groups)+
    stat_compare_means(comparisons=list(c("TB SM-", "LTBI SM-"), c("TB SM+", "LTBI SM+")),
                       method="wilcox.test", size=6)+
    coord_cartesian(ylim=c(0,35))+
    labs(y="(%) Total CD4 T cells",
        title="CFP-10 & ESAT-6")

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(seb, pep, nrow=1, rel_widths = c(3,2), labels=c("B", "C"), label_size = 20)
figure6a<-cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1.5,3))

second_row = cowplot::plot_grid(seb, pep2, nrow=1, rel_widths = c(3,2), labels=c("B", "C"), label_size = 20)
figure6b<-cowplot::plot_grid(first_row, second_row, ncol=1, rel_heights = c(1.5,3))

kable(table(og$disease), caption="Total Proliferation")
figure6a
figure6b
```

#Figure 7
##Mtb-Specific Responses in LTBI
```{r}
og<-filter_ltbi(dplyr::filter(OG, Stim=="PEP"))
ogphen<-filter_ltbi(dplyr::filter(OGPHEN, Stim=="PEP"))
oggt<-filter_ltbi(dplyr::filter(OG.GT, Stim=="PEP"))
ogth1<-filter_ltbi(dplyr::filter(OG.th1, Stim=="PEP"))

kable(table(og$disease), caption="PEP Proliferation")
kable(table(ogphen$disease[!is.na(ogphen$IFNg)]), caption="PEP phenotype")
```

####There are no differences in Mtb-specific proliferation betweeen SM+ and SM- QF+ individuals. The cytokine production from expanded Mtb-specific CD4 T cells only differs with regards to IFNγ IL-4 co-production, however the frequency of these cells is incredibly low. The dominant cytokine subgroup is still cells that co-produce IFNγ and TNFα. 
```{r og and cyt, fig.width=15, fig.height=8}
groups <- as_labeller(c("IFNg" = "IFNγ", 
                    "TNFa" = "TNFα", "IL4" = "IL-4"))

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
one<-baseplot(melt, "disease", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = groups, ncol=2)+
    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
two<-baseplot(melt, "disease", "value", ltbi.cols)+    
    facet_wrap(~variable, labeller = groups)
```

####Similar to the overnight assay, expanded Mtb-specific TH1 cytokine producing cells express both TH1 and TH2 lineage-specific markers, however this does not differ by S. mansoni status.

*I maybe want to do some analysis about the fact that now higher frequencies of cells express T-bet and/or GATA3 and that there seems to be a preferential expansion of CCR4+ cells but I honestly don't know how to test that*
###On proliferating cells
```{r og main four markers, fig.width=12, fig.height=12}
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3", "CXCR3", "CCR4"))
melt$variable<-factor(melt$variable, levels=c("T-bet", "GATA3", "CXCR3", "CCR4"))

tf<-baseplot(melt[melt$variable %in% c("T-bet","GATA3"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,100))+
    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")

sm<-baseplot(melt[melt$variable %in% c("CXCR3", "CCR4"),], "SM", "value", ltbi.cols)+    
    facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,100))

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20)
second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B"), label_size = 20)
third_row = cowplot::plot_grid(tf, sm, labels=c('C'), nrow=1, label_size = 20)
figure7<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3, rel_heights = c(2,2,2))
figure7
```

#Luminex Data

####Furthermore, IFNγ and TNFa are also the two dominant cytokines produced when analyzed by luminex. The concentration of all cytokines measured by luminex are equivalent between SM- and SM+ groups.

*Honestly I will have to just remake the whole thing a couple times and stitch them together to fix the axes so just ignore them for now*

```{r luminex, fig.width=12, fig.height=4}
lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
names(lum)<-gsub("IL", "IL-", names(lum))
names(lum)<-gsub("IFNg", "IFNγ", names(lum))
names(lum)<-gsub("TNFa", "TNFα", names(lum))

melt<-melt(lum, id.vars=c("TB", "SM"), measure.vars=grep("pg", names(lum)))
melt$variable<-gsub("_pg.ml", "", melt$variable)
melt$variable<- factor(melt$variable, levels=c("TNFα", "IFNγ", "IL-21", "IL-17A", "IL-22", "IL-10", "IL-4", "IL-5", "IL-13"))

a<-baseplot(melt[melt$variable=="TNFα",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,1500))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
b<-baseplot(melt[melt$variable=="IFNγ",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,18000))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
c<-baseplot(melt[melt$variable=="IL-21",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,60))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
d<-baseplot(melt[melt$variable=="IL-17A",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,20))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
e<-baseplot(melt[melt$variable=="IL-22",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,40))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
f<-baseplot(melt[melt$variable=="IL-10",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,20))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
g<-baseplot(melt[melt$variable=="IL-4",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,40))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
h<-baseplot(melt[melt$variable=="IL-5",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,15))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))
i<-baseplot(melt[melt$variable=="IL-13",], "SM", "value", ltbi.cols)+facet_wrap(~variable)+
    coord_cartesian(ylim=c(0,400))+
    theme(plot.margin=margin(l=-0.5,unit="cm"))

grid<-cowplot::plot_grid(a,b,c,d,e,f,g,h,i, nrow=1, align="hv")
lumplot<-cowplot::plot_grid(grid, NULL, ncol=1, align = "v", labels=c("A", "B"), rel_heights=c(3,1), label_size = 20)
lumplot
```

####However the correlation between cytokines differs between groups. Of note, the correlation between IFNγ and TNFa is stronger amongst SM+ individuals than SM- individuals.
```{r cor plots}
df<-lum[,3:11]
names(lum)<-gsub("_pg.ml", "", names(lum))
pca1 <- prcomp(df, scale = TRUE)
pcas<-data.frame(pca1$x)
rotations<-data.frame(pca1$rotation)
rotations$cytokine<-row.names(rotations)
melt<-melt(rotations, id.vars="cytokine")

stims<-factor(lum$Stim)
worms<-factor(lum$SM)
a<-ggplot(pcas, aes(x=pcas[,1], y=pcas[,2]))+geom_point(aes(col=worms))+labs(y="PCA2", x="PCA1")

neg <- cor(lum[lum$SM=="SM-",3:11], method="spearman")
order <- corrMatOrder(neg, order = "hclust")
order.neg <- neg[order, order]
res1 <- cor.mtest(lum[lum$SM=="SM-",3:11], conf.level = .95)
pvals <- res1$p[order, order]
corrplot(order.neg, method = "circle", insig = "label_sig", tl.col = "black", 
         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
         type="upper")

pos <- cor(lum[lum$SM=="SM+",3:11], method="spearman")
order.pos <- pos[order, order]
res1 <- cor.mtest(lum[lum$SM=="SM+",3:11], conf.level = .95)
pvals <- res1$p[order, order]
corrplot(order.pos, method = "circle", insig = "label_sig", tl.col = "black", 
         p.mat = pvals, sig.level = c(.001, .01, .05), pch.cex = 2, pch.col = "white",
         type="upper")
```

#Exporting Data
```{r}
setwd("/Users/tarynam/Desktop/Manuscript1")

cowplot::save_plot("figure1.png", figure1, base_height=8, base_width=16) #Figure 1
#cowplot::save_plot("figure1alt1.png", figure1B, base_height=8, base_width=16) #Figure 1
cowplot::save_plot("figure1alt2.png", figure1alt, base_height=8, base_width=16) #Figure 1
cowplot::save_plot("figure2.png", figure2, base_height=14, base_width=16) #Figure 2
cowplot::save_plot("figure3.png", figure3, base_height=16, base_width=16)

cowplot::save_plot("figure4.png", figure4opt1, base_height=10, base_width=16)

cowplot::save_plot("figure5.png", figure5, base_height=11, base_width=16)
cowplot::save_plot("figure6a.png", figure6a, base_height=9, base_width=9)
cowplot::save_plot("figure6b.png", figure6b, base_height=9, base_width=9)
cowplot::save_plot("figure7.png", figure7, base_height=14, base_width=12)

cowplot::save_plot("figure8.png", lumplot, base_height=6, base_width=18)
```