### 2019-10-03 SuppFigures.Rmd

---
title: "supplementary figures"
author: "Taryn McLaughlin"
date: "9/26/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, comment=NA, warning=FALSE, echo=FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
library(tidyr)
library(reshape2)
library(cowplot)
library(corrplot)
library(sjPlot)
library(ggsignif)
```

```{r plot functions}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/comp_hc_ltbi.R")
```

```{r load data}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")
CD4<-read.csv("ICS_CD4_clean.csv")
CYT<-read.csv("ICS_CD4_Cytokine_clean.csv")
PHEN<-read.csv("ICS_CD4_phenotyping_clean_NAs.csv")
OG<-read.csv("OG_CD4_bulk_summary_clean.csv")
OGPHEN<-read.csv("OG_CD4_OG_summary_clean_NAs.csv")
OGnope<-read.csv("OG_CD4_noOG_summary_clean.csv")
OG.IFNg<-read.csv("OG_CD4_OG.IFNg_summary_clean_NAs.csv")
OG.GT<-read.csv("OG_CD4_OG.GT_summary_clean_NAs.csv")
OG.th1<-read.csv("OG_CD4_OG.TH1_summary_clean_NAs.csv")
luminex<-read.csv("/Applications/Old Computer/Day Lab/Luminex/Clean Data/Luminex_combined_background.txt")
```

```{r, all the labels I use multiple times}
#cytokines
G   <-c("+", "+", "+", "+", "-", "+", "+", "-", "-", "+", "+", "-", "-", "-", "-")
il4 <-c("+", "+", "-", "+", "+", "+", "-", "+", "-", "-", "-", "-", "+", "+", "-")
il13<-c("+", "-", "+", "+", "+", "-", "+", "-", "+", "-", "-", "-", "+", "-", "+")
T   <-c("+", "+", "+", "-", "+", "-", "-", "+", "+", "+", "-", "+", "-", "-", "-")

cyt_labels <- paste0(G, "\n", T, "\n", il4, "\n", il13)
cyt_breaks <- c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x")

#TH groups
th_groups <- as_labeller(c("TH1.Freq" = "TH1", 
                    "TH1.2.Freq" = "TH1/2", "TH2.Freq" = "TH2"))

#Phenotype boolean labels
T <- c("+", "+","+","+","+","+","+","+","-","-","-","-","-","-","-")
G <- c("-","-","-","-","+", "+","+","+","+","+","+","+","-","-","-")
c3 <-c("+", "+","-","-","+", "+","-","-","+", "+","-","-","+", "+","-")
c4 <-c("+","-","+","-","+","-","+","-","+","-","+","-","+","-","+")

phen_labels <- paste0(T, "\n", G, "\n", c3, "\n", c4)
phen_breaks <- c(
"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
"x_x_3_4" , "x_x_3_x" , "x_x_x_4")
```

#ICS Supplementary

##2 Boolean PMA
```{r}
cyt<-plot_filter(CYT[CYT$Stim=="PMA",])
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

th1.2<-booleanplot2(test[test$th=="TH1/2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,3))+
    theme(strip.text.y = element_blank(), strip.background.y = element_blank())+
    labs(y="(%) CD4 T cells")

th1<-booleanplot2(test[test$th=="TH1",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,25))+
    theme(strip.text.y = element_blank(), strip.background.y = element_blank())

th2<-booleanplot2(test[test$th=="TH2",], "variable", "value", all.cols)+facet_grid(TB~th, scales="free", space="free")+
    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 20)+
    coord_cartesian(ylim=c(0,3))
```

```{r, fig.height=10, fig.width=10}
supp2<-cowplot::plot_grid(th1.2, th1, th2, ncol=3, rel_widths = c(8,3.5,3.5))
supp2
```

##3 PMA Phenotypes
```{r}
phen<-plot_filter(subset(PHEN, PHEN$Stim=="PMA"))

test<-melt(phen, id.vars=c("TB", "SM", "disease"), measure.vars = grep("TH._._._._.", names(phen)))
test$thgroup<-"TH1"
test$thgroup[grep("TH2", test$variable)]<-"TH2"

test<-test[-(grep("x_x_x_x", test$variable)),]
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-gsub("TH2_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

supp3a<-booleanplot2(test[test$thgroup=="TH1",], "variable", "value", all.cols)+
    facet_grid(TB~tf, scales="free_x")+
    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,65))+
    theme(plot.margin=margin(l=1,unit="cm"))+
    labs(y="(%) TH1 Cytokine+ CD4 T cells")

supp3b<-booleanplot2(test[test$thgroup=="TH2",], "variable", "value", all.cols)+
    facet_grid(TB~tf, scales="free_x")+
    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,65))+
    theme(plot.margin=margin(l=1,unit="cm"))+
    labs(y="(%) TH2 Cytokine+ CD4 T cells")

```

```{r, fig.height=12, fig.width=10}
supp3<-cowplot::plot_grid(supp3a, supp3b, nrow=2, labels=c("A", "B"), label_size = 20)
supp3
```

##4 LTBI ICS extra 
```{r}
cyt<-filter_ltbi(CYT[CYT$Stim=="PEP",])

#boolean cytokine
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

a<-booleanplot2(test, "variable", "value", all.cols)+facet_grid(~th, scales="free", space="free")+
    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
    coord_cartesian(ylim=c(0,.16))+
    theme(plot.margin=margin(l=1,unit="cm"))+
    labs(y="(%) CD4 T cells")

#condensed TH cytokine
test<-melt(cyt,
           id.vars=c("TB","SM"), 
           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))

b<-baseplot(test, "SM", "value", ltbi.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,.2))+
    labs(y="(%) CD4 T cells")

#phenotypes
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))

melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

c<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
    theme(legend.position = "right", legend.key.size = unit(4, "cm"))+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10)+
    theme(plot.margin=margin(l=0.5,unit="cm"))+
    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")
```

```{r, fig.height=10, fig.width=12}
#compiling together
suppab<-cowplot::plot_grid(a, b, nrow=1, labels=c("A", "B"), label_size = 20,
                           align="hv", rel_widths = c(6,3))
suppc<-cowplot::plot_grid(c, nrow=1, labels=c("C"), label_size = 20)
supp4<-cowplot::plot_grid(suppab, suppc, align="hv", nrow=2)
supp4
```

##5 Comparing Mtb-specific CD4 T cells to other CD4 T cells in LTBI
```{r}
total<-filter_ltbi(subset(CD4, CD4$Stim=="UN"))
total$cell<-"Total"
phen<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
mtb_th1<-select(phen, names)
names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
mtb_th1$cell<-"Mtb-Specific"
long<-rbindlist(list(total, mtb_th1), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
melt<-data.frame(melt)

a1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
    labs(y="(%) CD4 T cells")+
    geom_signif(
        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*32)+    
    coord_cartesian(ylim=c(0,32))

a2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
    geom_signif(
        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 98)+    
    coord_cartesian(ylim=c(0,100))
```

```{r}
pma<-filter_ltbi(subset(PHEN, PHEN$Stim=="PMA"))
pma$cell<-"PMA"
pep<-filter_ltbi(subset(PHEN, PHEN$Stim=="PEP"))
pep$cell<-"Mtb"
long<-rbindlist(list(pma, pep), fill=TRUE)
names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
long<-select(long, names)
names(long)<-gsub("TH1_", "", names(long))

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
melt<-data.frame(melt)

b1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
    labs(y="(%) CD4 T cells")+
    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*40)+    
    coord_cartesian(ylim=c(0,40))

b2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 98)+    
    coord_cartesian(ylim=c(0,100))
```

```{r, fig.height=10, fig.width=12}
supp5<-cowplot::plot_grid(a1,a2,b1,b2, ncol=2, labels=c("A","","B",""), label_size = 20)
supp5
```

##6 TB ICS extra
```{r}
cyt<-filter_tb(CYT[CYT$Stim=="PEP",])

#boolean cytokine
test<-melt(cyt, id.vars=c("TB", "SM", "disease"), measure.vars = grep("._._._.|._._.._.", names(cyt)))
test<-test[-grep("tbet|gata3|x_x_x_x", test$variable),]
test$variable<-factor(test$variable, levels=c(
"G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", "G_x_13_x","x_4_x_T","x_x_13_T",
"G_x_x_T" , "G_x_x_x","x_x_x_T",
"x_4_13_x","x_4_x_x", "x_x_13_x"
))

test$th<-"TH1/2"
test$th[grep("._x_x_.", test$variable)]<-"TH1"
test$th[grep("x_._._x|x_._.._x", test$variable)]<-"TH2"
test$th<-factor(test$th, levels=c("TH1/2", "TH1", "TH2"))

a<-booleanplot2(test, "variable", "value", all.cols)+facet_grid(~th, scales="free", space="free")+
    scale_x_discrete(labels = cyt_labels, breaks=cyt_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, exact=FALSE, size=10)+
    coord_cartesian(ylim=c(0,.16))+
    theme(plot.margin=margin(l=1,unit="cm"))+
    labs(y="(%) CD4 T cells")

#condensed TH cytokine
test<-melt(cyt,
           id.vars=c("TB","SM"), 
           measure.vars=c("TH1.Freq", "TH1.2.Freq", "TH2.Freq"))

b<-baseplot(test, "SM", "value", tb.cols)+
    facet_wrap(~variable, ncol=4, labeller=th_groups)+
    coord_cartesian(ylim=c(0,.2))+
    labs(y="(%) CD4 T cells")

#phenotypes
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))

melt<-melt(phen, id.vars=c("TB", "SM"), measure.vars = grep("TH1_._._._.", names(phen)))
test<-filter(melt, variable!="TH1_x_x_x_x")
test$variable<-gsub("TH1_", "", test$variable)
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

c<-booleanplot(test, "variable", "value", tb.cols)+facet_wrap(~tf, scales="free_x",ncol=4)+
    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
    theme(legend.position = "right", legend.key.size = unit(4, "cm"))+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=10, label.y = 40)+
    coord_cartesian(ylim=c(0,50))+
    labs(y="(%) Mtb-Specific\nTH1 Cytokine+ cells")
```

```{r, fig.height=10, fig.width=12}
#compiling together
suppab<-cowplot::plot_grid(a, b, nrow=1, labels=c("A", "B"), label_size = 20, 
                           align="hv", rel_widths = c(6,3))
suppc<-cowplot::plot_grid(c, nrow=1, labels=c("C"), label_size = 20)
supp6<-cowplot::plot_grid(suppab, suppc, align="v", nrow=2)
supp6
```

##7 Comparing Mtb-specific CD4 T cells to others in TB
```{r}
total<-filter_tb(subset(CD4, CD4$Stim=="UN"))
total$cell<-"Total"
phen<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
names<-c("Donor", "TB", "SM", "Stim", "disease", names(phen)[grep("TH1_", names(phen))])
mtb_th1<-select(phen, names)
names(mtb_th1)<-gsub("TH1_", "", names(mtb_th1))
mtb_th1$cell<-"Mtb-Specific"
long<-rbindlist(list(total, mtb_th1), fill=TRUE)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$cell<-factor(melt$cell, levels=c("Total", "Mtb-Specific"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- Total", "SM- Mtb-Specific","SM+ Total", "SM+ Mtb-Specific"))
melt<-data.frame(melt)

a1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", tb.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
    labs(y="(%) CD4 T cells")+
    geom_signif(
        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
    coord_cartesian(ylim=c(0,80))

a2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", tb.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("Total", "CFP-10 &\nESAT-6", "Total", "CFP-10 &\nESAT-6"))+
    geom_signif(
        comparisons=list(c("SM- Mtb-Specific", "SM- Total"),c("SM+ Mtb-Specific", "SM+ Total")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
    coord_cartesian(ylim=c(0,80))

```

```{r}
pma<-filter_tb(subset(PHEN, PHEN$Stim=="PMA"))
pma$cell<-"PMA"
pep<-filter_tb(subset(PHEN, PHEN$Stim=="PEP"))
pep$cell<-"Mtb"
long<-rbindlist(list(pma, pep), fill=TRUE)
names<-c("Donor", "TB", "SM", "Stim", "disease", "cell", names(long)[grep("TH1_", names(long))])
long<-select(long, names)
names(long)<-gsub("TH1_", "", names(long))

melt<-melt(long, id.vars=c("TB", "SM", "disease", "cell"), measure.vars=c("T-bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell, sep=" ")
melt$x<-factor(melt$x, levels=c("SM- PMA", "SM- Mtb","SM+ PMA", "SM+ Mtb"))
melt<-data.frame(melt)

b1<-baseplot(melt[(melt$variable %in% c("T-bet", "GATA3")),], "x", "value", tb.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
    labs(y="(%) CD4 T cells")+
    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
    coord_cartesian(ylim=c(0,80))

b2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", tb.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("PMA", "CFP-10 &\nESAT-6", "PMA", "CFP-10 &\nESAT-6"))+
    geom_signif(comparisons=list(c("SM- Mtb", "SM- PMA"),c("SM+ Mtb", "SM+ PMA")), 
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = .98*80)+    
    coord_cartesian(ylim=c(0,80))
```

```{r, fig.height=10, fig.width=12}
supp7<-cowplot::plot_grid(a1,a2,b1,b2, ncol=2, labels=c("A","","B",""), label_size = 20)
supp7
```

#Proliferation Supplementary

##8 SEB
```{r}
ogphen<-plot_filter(subset(OGPHEN, OGPHEN$Stim=="SEB"))
groups <- as_labeller(c("OG" = "Proliferation","IFNg" = "IFNγ","TNFa" = "TNFα", "IL4" = "IL-4"))
#cytokines
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IFNg", "TNFa"))
one<-allplot(melt, "TB", "value", all.cols)+    
    facet_wrap(~variable, labeller = groups, ncol=2)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="(%) Proliferating\nCD4 T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("IL4"))
two<-allplot(melt, "TB", "value", all.cols)+    
    facet_wrap(~variable, labeller = groups)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    coord_cartesian(ylim=c(0,5))

#phenotypes
melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("T-bet", "GATA3"))
tf<-allplot(melt, "TB", "value", all.cols)+    
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))+
    labs(y="(%) Proliferating\nCD4 T cells")

melt<-melt(ogphen, id.vars=c("TB","SM", "disease"), measure.vars=c("CXCR3", "CCR4"))
sm<-allplot(melt, "TB", "value", all.cols)+
    scale_x_discrete(labels=c("SM- SM+\nHC", "SM- SM+\nLTBI", "SM- SM+\nTB"))

first_row = cowplot::plot_grid(NULL, labels = c('A'), nrow=1, label_size = 20, rel_widths = c(3,1))
second_row = cowplot::plot_grid(one, two, nrow=1, rel_widths = c(2,1), labels=c("B"), label_size = 20)
third_row= cowplot::plot_grid(tf, sm, labels=c('C'), nrow=1, label_size = 20)
```

```{r, fig.height=12, fig.width=10}
supp8<-cowplot::plot_grid(first_row, second_row, third_row, nrow=3)
supp8
```

##9 LTBI Peptide
```{r}
ogphen<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))

#Cytokine
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._.", names(ogphen)))
test<-melt[-grep("._._._.", melt$variable),]
test<-filter(test, variable!="x_x_x")
test$variable<-factor(test$variable, levels=c(
"T_G_x" , "T_x_x" , "x_G_x" ,         
"T_G_4" , "T_x_4" , "x_G_4" , 
"x_x_4"))

test$group<-"TH1/2"
for(i in 1:length(test$group)){
    if(test$variable[i] %in% c("T_G_x", "T_x_x", "x_G_x")){
        test$group[i]<- "TH1"
    }
    else if (test$variable[i] == "x_x_4"){
        test$group[i]<- "TH2"
    }
}

T <- c("+", "+", "-", "+", "+", "-", "-")
G <- c("+", "-", "+", "+", "-", "+", "-")
il4 <- c("-", "-", "-", "+", "+", "+", "+")

labels <- paste0(T, "\n", G, "\n", il4)
breaks <- levels(test$variable)

a<-booleanplot(test, "variable", "value", ltbi.cols)+
    scale_x_discrete(labels = labels, breaks=breaks)+
    theme(legend.position = "right", legend.key.size = unit(6, "cm"))+
    facet_grid(~group, scales="free_x", space="free_x")+
    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")

#phenotype
melt<-melt(ogphen, id.vars=c("TB", "SM"), measure.vars = grep("._._._.", names(ogphen)))
test<-filter(melt, variable!="x_x_x_x")
test$variable<-factor(test$variable, levels=c(
"t_x_3_4" , "t_x_3_x" , "t_x_x_4" , "t_x_x_x" ,
"t_g_3_4" , "t_g_3_x" , "t_g_x_4" , "t_g_x_x" ,
"x_g_3_4" , "x_g_3_x" , "x_g_x_4" , "x_g_x_x" ,
"x_x_3_4" , "x_x_3_x" , "x_x_x_4"))

test$tf<-NA
test$tf[grep("x_g", test$variable)]<-"GATA3"
test$tf[grep("t_x", test$variable)]<-"T-bet"
test$tf[grep("t_g", test$variable)]<-"T-bet & GATA3"
test$tf[grep("x_x_._.", test$variable)]<-"None"
test$tf<-factor(test$tf, levels=c("T-bet", "T-bet & GATA3", "GATA3", "None"))

b<-booleanplot(test, "variable", "value", ltbi.cols)+facet_wrap(~tf, scales="free_x", ncol=4)+
    scale_x_discrete(labels = phen_labels, breaks=phen_breaks)+
    stat_compare_means(method="wilcox.test", label="p.signif", hide.ns = TRUE, size=12)+
    labs(y="(%) Proliferating\nMtb-Specific CD4 T cells")
```

```{r, fig.height=10, fig.width=10}
supp9<-cowplot::plot_grid(a, b, nrow=2, labels=c("A", "B"), label_size = 20, rel_widths = c(9,10))
supp9
```

##10 CFP-10 and ESAT-6-proliferating to other cells
```{r}
un<-filter_ltbi(dplyr::filter(OGnope, Stim=="PEP"))
un$cell<-"Non-Proliferating"
ce<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
ce$cell<-"Proliferating"
long<-rbindlist(list(un, ce), fill=TRUE)
long$cell<-factor(long$cell, levels=c("Non-Proliferating", "Proliferating"))
long<-data.frame(long)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell)
melt$x<-factor(melt$x, levels=c("SM- Non-Proliferating", "SM- Proliferating", 
                                "SM+ Non-Proliferating", "SM+ Proliferating"))

a1<-baseplot(melt[(melt$variable %in% c("T.bet", "GATA3")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("NP", "P", 
                              "NP", "P"))+
    geom_signif(comparisons=list(c("SM- Non-Proliferating", "SM- Proliferating"),
                                 c("SM+ Non-Proliferating", "SM+ Proliferating")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 100)+
    labs(y="(%) CD4 T cells")+
    coord_cartesian(ylim=c(0,105))

a2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("NP", "P", 
                              "NP", "P"))+
    geom_signif(comparisons=list(c("SM- Non-Proliferating", "SM- Proliferating"),
                                 c("SM+ Non-Proliferating", "SM+ Proliferating")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 8)+
    coord_cartesian(ylim=c(0,105))

```

```{r}
un<-filter_ltbi(dplyr::filter(OGPHEN, Stim=="SEB"))
un$cell<-"SEB"
ce<-filter_ltbi(subset(OGPHEN, OGPHEN$Stim=="PEP"))
ce$cell<-"PEP"
long<-rbindlist(list(un, ce), fill=TRUE)
long$cell<-factor(long$cell, levels=c("SEB", "PEP"))
long<-data.frame(long)

melt<-melt(long, id.vars=c("TB", "SM", "disease", "Donor", "cell"), 
           measure.vars=c("T.bet", "GATA3", "CXCR3","CCR4"))
melt$x<-paste(melt$SM, melt$cell)
melt$x<-factor(melt$x, levels=c("SM- SEB", "SM- PEP", 
                                "SM+ SEB", "SM+ PEP"))

b1<-baseplot(melt[(melt$variable %in% c("T.bet", "GATA3")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("SEB", "CFP-10 &\nESAT-6", "SEB", "CFP-10 &\nESAT-6"))+
    labs(y="(%) CD4 T cells")+
    geom_signif(comparisons=list(c("SM- SEB", "SM- PEP"),c("SM+ SEB", "SM+ PEP")),
                test = "wilcox.test", map_signif_level=TRUE, textsize = 6, tip_length = 0.03, y_position = 100)+
    coord_cartesian(ylim=c(0,105))

#not significant so didn't include
b2<-baseplot(melt[(melt$variable %in% c("CXCR3", "CCR4")),], "x", "value", ltbi.cols)+
    facet_wrap(~variable, nrow=1)+
    scale_x_discrete(labels=c("SEB", "CFP-10 &\nESAT-6", "SEB", "CFP-10 &\nESAT-6"))+
    coord_cartesian(ylim=c(0,100))
```

```{r, fig.height=10, fig.width=12}
supp10<-cowplot::plot_grid(a1,a2, b1,b2, ncol=2, labels=c("A", "","B", ""), label_size = 20)
supp10
```

##11 Dendrograms
```{r}
lum<-filter_ltbi(subset(luminex, luminex$Stim=="PEP"))
names(lum)<-gsub("IL", "IL-", names(lum))
names(lum)<-gsub("IFNg", "IFNγ", names(lum))
names(lum)<-gsub("TNFa", "TNFα", names(lum))

lum_x<-lum[lum$SM=="SM-",]
df_x<-as.matrix(lum_x[,3:11])
row.names(df_x)<-lum_x$Sample
hc_x <- hclust(as.dist(1-cor(df_x)))
x<-plot(hc_x)

lum_sm<-lum[lum$SM=="SM+",]
df_sm<-as.matrix(lum_sm[,3:11])
row.names(df_sm)<-lum_sm$Sample
hc_sm <- hclust(as.dist(1-cor(df_sm)))
sm<-plot(hc_sm)

df<-as.matrix(lum[,3:11])
#heatmap(t(df_x), Rowv=as.dendrogram(hc_x))
#heatmap(t(df_sm), Rowv=as.dendrogram(hc_sm))
#heatmap(t(df), Rowv=as.dendrogram(hc_sm))
```

#Exporting Data
```{r}
setwd("/Users/tarynam/Desktop/Manuscript1")
cowplot::save_plot("supp2.png", supp2, base_height=8, base_width=10)
cowplot::save_plot("supp3.png", supp3, base_height=14, base_width=10)
cowplot::save_plot("supp4.png", supp4, base_height=12, base_width=12)
cowplot::save_plot("supp5.png", supp5, base_height=15, base_width=16)
cowplot::save_plot("supp6.png", supp6, base_height=12, base_width=12)
cowplot::save_plot("supp7.png", supp7, base_height=15, base_width=16)
cowplot::save_plot("supp8.png", supp8, base_height=14, base_width=16)
cowplot::save_plot("supp9.png", supp9, base_height=14, base_width=16)
cowplot::save_plot("supp10.png", supp10, base_height=15, base_width=16)
```

#Extra plots

To see if the variation in T-bet and GATA3 in the SEB proliferation was due to original variation or proliferation. It's not.
```{r}
test<-merge(OG, OGPHEN, by=c("Sample", "Donor", "Stim", "TB", "SM"))
test<-filter(test, Stim=="SEB")
ggplot(test, aes(x=OG.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")+
    labs(x="Proliferation (%) CD4 T cells",
         y="T-bet (%) Proliferating CD4 T cells")
ggplot(test, aes(x=OG.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
        labs(x="Proliferation (%) CD4 T cells",
         y="GATA3 (%) Proliferating CD4 T cells")
ggplot(test, aes(x=Tbet.y, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
    labs(x="T-bet (%) Proliferating CD4 T cells",
         y="GATA3 (%) Proliferating CD4 T cells")

part1<-filter(OGPHEN, Stim=="UN")
part2<-filter(OGPHEN, Stim=="SEB")
test<-merge(part1, part2, by=c("Donor", "TB", "SM"))
ggplot(test, aes(x=Tbet.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")+
        labs(x="(%) CD4 T cells",
         y="(%) Proliferating CD4 T cells",
         title="T-bet")
ggplot(test, aes(x=GATA3.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
        labs(x="(%) CD4 T cells",
         y="(%) Proliferating CD4 T cells",
         title="GATA3")

part1<-filter(CD4, Stim=="UN")
part2<-filter(OGPHEN, Stim=="SEB")
test<-merge(part1, part2, by=c("Donor", "TB", "SM"))
ggplot(test, aes(x=Tbet.x, y=Tbet.y, color=TB))+geom_point()+stat_smooth(method="lm")+
        labs(x="(%) ICS CD4 T cells",
         y="(%) Proliferating CD4 T cells",
         title="T-bet")
ggplot(test, aes(x=GATA3.x, y=GATA3.y, color=TB))+geom_point()+stat_smooth(method="lm")+
        labs(x="(%) ICS CD4 T cells",
         y="(%) Proliferating CD4 T cells",
         title="GATA3")
```